### 2023-06-14 app/melding_link_to_page.Rmd
initialize

---
title: "Bayesian Melding App"
author: "Quinn White"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    df_print: paged
    code_folding: hide
    css: ../css/template.css
  html_document:
    df_print: paged
---

```{css, echo = FALSE}
#content{
max-width:1920px;
}

```

Shiny app exploring the implications of changing the parameters for $\alpha, \beta,$ and $P(S_1|\text{untested})$ is deployed [here](https://q-w-a.shinyapps.io/bayesian_melding_priors/).

<br><br><br><br><br><br><br><br><br>