### 2023-06-19 reports/county_correlations.Rmd
update report files

---
title: "County Correlations"
author: "Quinn White"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    df_print: paged
    code_folding: hide
    css: !expr here::here('css/template.css')
  html_document:
    df_print: paged
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = "../output_html")})
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE,warning=FALSE)

library(tidyverse)
library(targets)
library(here)
library(latex2exp)

```

```{r theme_c}

theme_c <- function(...){ 
   # font <- "Helvetica"   #assign font family up front
    theme_bw() %+replace%    #replace elements we want to change

    theme(

      #text elements
      plot.title = element_text(             #title
                   size = 16,                #set font size
                   face = 'bold',            #bold typeface
                   hjust = .5,
                   vjust = 3),

      plot.subtitle = element_text(          #subtitle
                   size = 12,
                   hjust = .5,
                   face = 'italic',
                   vjust = 3),               #font size

      axis.title = element_text(             #axis titles
                   size = 12),               #font size

      axis.text = element_text(              #axis text
                   size = 9),
      legend.text = element_text(size = 12),
      # t, r, b, l
      plot.margin = unit(c(1,.5,.5,.5), "cm"),
      legend.position = "right",
      strip.text.x = element_text(size = 11, face = "bold", color="white"),
      strip.background = element_rect(fill = "#3E3D3D")
      ) %+replace%
      theme(...)

}

```

```{r load-data}

county <- tar_read(ma_v1, 
                   store = here("_targets")) %>% 
  mutate(version = "v1") %>%
  bind_rows(
    tar_read(ma_v2,
             store = here("_targets"))
  ) %>%
   bind_rows(
    tar_read(ma_v3, 
             store = here("_targets"))
  ) %>%
   bind_rows(
    tar_read(ma_v4, 
             store = here("_targets"))
  )

waste <- tar_read(waste, store = here("_targets"))
covidestim <- tar_read(covidestim_biweekly_county,
                        store = here("_targets"))

```

```{r}

wjoined <- county %>%
  # set Nantucket, Duke fips to Nantucket since we have wastewater data
  # for Nantucket
  mutate(fips = ifelse(grepl("25019", fips), "25019", fips)) %>%
  inner_join(waste) %>%
  left_join(covidestim, relationship = "many-to-many") %>%
  group_by(biweek) %>%
  mutate(mindate = min(date),
         maxdate = max(date)) %>%
  ungroup()

counties_with_waste <- wjoined %>%
  group_by(fips) %>%
  mutate(obs_w_notna = sum(!is.na(mean_conc))) %>%
  filter(obs_w_notna != 0) %>%
  pull(fips) %>%
  unique()

versions <- tibble(
  version = c("v1", "v2", "v3", "v4"),
  vlabel = factor(
    c( "$Priors\\,Do\\,Not\\,Vary\\,by\\,County\\,and\\,Date$", 
    "$\\beta$ Centered at Emp. Value",
    "$P(S_1|untested)$ Centered at Emp. Value",
    "$P(S_1|untested)$ and $\\beta$ Centered at Emp. Values"
    ),
    levels = c(
        "$Priors\\,Do\\,Not\\,Vary\\,by\\,County\\,and\\,Date$", 
        "$\\beta$ Centered at Emp. Value",
         "$P(S_1|untested)$ Centered at Emp. Value",
        "$P(S_1|untested)$ and $\\beta$ Centered at Emp. Values"

    ) )
)

wjoined <- wjoined %>%
  left_join(versions)

w_data <- wjoined %>% 
  select(-date) %>%
  distinct() %>%
  filter(fips %in% counties_with_waste) %>%
  filter(!is.na(mean_conc)) %>%
  group_by(fips) %>%
  mutate(mean_conc_std = mean_conc/ max(mean_conc),
         exp_cases_median_std = exp_cases_median / max(exp_cases_median),
         exp_cases_lb_std = exp_cases_lb /max(exp_cases_median),
         exp_cases_ub_std = exp_cases_ub / max(exp_cases_median),
         infections_std  = infections/max(infections),
         positive_std = positive/max(positive)) %>%
  ungroup()

```

# Correlations

For each county, normalize the wastewater concentration by the maximum concentration for that county and the probabilistic bias counts by the maximum *median* estimated counts for that county (2.5th percentile and 97.5th percentile also standardized by the maximum *median* estimated counts for that county).

## Wastewater

```{r}

# correlations between PB counts and wastewater concentrations

correlations <- w_data %>%
  group_by(version) %>%
  summarize(Correlation = cor(exp_cases_median_std,mean_conc_std),
            vlabel =unique(vlabel)) %>%
  mutate(x=1.3,
         y=.1)

w_data %>%
  group_by(version) %>%
  ggplot(aes(x=exp_cases_median_std, y = mean_conc_std)) +
  geom_point(alpha=.5) +
  geom_linerange(aes(xmin = exp_cases_lb_std,
                     xmax=exp_cases_ub_std, 
                     y = mean_conc_std),
                 alpha = .3) +
  geom_text(
           aes(x=x,y=y,  label=paste0("Correlation: ", 
                                      round(Correlation,3))), 
                data= correlations ) +
  facet_wrap(~vlabel,
             labeller= as_labeller(TeX, 
                                   default = label_parsed)) +
  theme_c() +
  labs(x = "Probabilistic Bias Counts\n(Standardized)",
       y = "Mean Wastewater Concentration\n(Standardized)",
       title = "Correlations Between Probabilistic Bias Counts\nand Mean Wastewater Concentration")

```

## Covidestim

```{r}

# correlations between PB counts and Covidestim estimates

covidestim_correlations <- w_data %>%
  filter(!is.na(infections_std)) %>%
  group_by(vlabel) %>%
  # missing values are handled by casewise deletion
  summarize(Correlation = cor(exp_cases_median_std,infections_std,
                              use = "complete.obs")) %>%
  mutate(x=1.3,
         y=.1)

w_data %>%
  ggplot(aes(x=exp_cases_median_std, y = infections_std)) +
  geom_point(alpha=.5) +
  geom_linerange(aes(xmin = exp_cases_lb_std,
                     xmax=exp_cases_ub_std, 
                     y = infections_std),
                 alpha = .3) +
  geom_text(
           aes(x=x,y=y,  label=paste0("Correlation: ", 
                                      round(Correlation,3))), 
                data= covidestim_correlations ) +
  facet_wrap(~vlabel,
             labeller= as_labeller(TeX, 
                                   default = label_parsed)) +
  theme_c() +
  labs(x = "Probabilistic Bias Counts\n(Standardized)",
       y = "Covidestim Estimates\n(Standardized)",
       title = "Correlations Between Probabilistic Bias Counts\nand Covidestim Estimates")

```

## Covidestim and Wastewater

```{r}

covidestim_correlations <- w_data %>%
  filter(version=="v1") %>%
  # missing values are handled by casewise deletion
  summarize(Correlation = cor(mean_conc_std,infections_std,
                              use = "complete.obs")) %>%
  mutate(x=1.3,
         y=.1)

w_data %>%
  filter(version=="v1") %>%
  ggplot(aes(x=infections_std, y = mean_conc_std)) +
  geom_point(alpha=.5) +
  geom_text(
           aes(x=x,y=y,  label=paste0("Correlation: ", 
                                      round(Correlation,3))), 
                data= covidestim_correlations ) +
  facet_wrap(~vlabel,
             labeller= as_labeller(TeX, 
                                   default = label_parsed)) +
  theme_c() +
  labs(x = "Covidestim Estimates\n(Standardized)",
       y = "Wastewater Concentrations\n(Standardized)",
       title = "Correlations Between Probabilistic Bias Counts\nand Covidestim Estimates")

pb_correlations <- w_data %>%
  filter(version=="v1") %>%
  # missing values are handled by casewise deletion
  summarize(Correlation = cor(positive_std,mean_conc_std,
                              use = "complete.obs")) %>%
  mutate(x=1.3,
         y=.1)

```