### 2023-06-20 reports/priors.Rmd

---
title: "Priors"
author: "Quinn White"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    df_print: paged
    code_folding: hide
    css: !expr here::here('css/template.css')
  html_document:
    df_print: paged
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = "../output_html")})
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)

library(tidyverse)
library(targets)
library(here)
library(latex2exp)

# read priors from target
prior_params <- tar_read(prior_params,
                         store = here("_targets"))
# load beta_density, gamma_density functions
source(here("R/03-base-functions.R"))

theme_c <- function(...){ 
   # font <- "Helvetica"   #assign font family up front
    theme_bw() %+replace%    #replace elements we want to change

    theme(

      #text elements
      plot.title = element_text(             #title
                   size = 16,                #set font size
                   face = 'bold',            #bold typeface
                   hjust = .5,
                   vjust = 3),

      plot.subtitle = element_text(          #subtitle
                   size = 12,
                   hjust = .5,
                   face = 'italic',
                   vjust = 3),               #font size

      axis.title = element_text(             #axis titles
                   size = 12),               #font size

      axis.text.x = element_text(              #axis text
                   size = 12),
      legend.text = element_text(size = 14),
      # t, r, b, l
      plot.margin = unit(c(1,.5,.5,.5), "cm"),
      legend.position = "right",
      strip.text.x = element_text(size = 16, face = "bold", color="white"),
      strip.background = element_rect(fill = "#3E3D3D")
      ) %+replace%
      theme(...)

}

```

# $\alpha$

```{r}

tibble(x = seq(0, 1.3, length = 10^(5)),
       y = with(prior_params, gamma_density(x,
                                     mean = alpha_mean,
                                     sd = alpha_sd,
                                     bounds = alpha_bounds))) %>%
   ggplot(aes(x=x, y = y)) +
  geom_line(alpha = .8) + 
  geom_ribbon(aes(x=x,ymin=0,ymax=y),
              fill="black",
              alpha=.6) +
  theme_c(legend.text=element_text(size = 10)) +
  viridis::scale_fill_viridis(discrete=TRUE,
                             option = "rocket", begin=.3,end=.8) +
  labs(x = "Value",
       y = "Probability Density", 
       title = latex2exp::TeX("Definition of Prior for $\\alpha$"),
       fill ='',
       subtitle = paste0("Mean: ", prior_params$alpha_mean,
                         ", SD: ", prior_params$alpha_sd))

```

# $\beta$

```{r}

tibble(x = seq(0, 1, length = 10^(5)),
       y = with(prior_params, beta_density(x,
                                     mean = beta_mean,
                                     sd = beta_sd,
                                     bounds = beta_bounds))) %>%
   ggplot(aes(x=x, y = y)) +
  geom_line(alpha = .8) + 
  geom_ribbon(aes(x=x,ymin=0,ymax=y),
              fill="black",
              alpha=.6) +
  theme_c(legend.text=element_text(size = 10)) +
  viridis::scale_fill_viridis(discrete=TRUE,
                             option = "rocket", begin=.3,end=.8) +
  labs(x = "Value",
       y = "Probability Density", 
       title = latex2exp::TeX("Definition of Prior for $\\beta$"),
       fill ='',
       subtitle = paste0("Mean: ", prior_params$beta_mean,
                         ", SD: ", prior_params$beta_sd))

```

# $P(S_1|\text{untested})$

```{r}

tibble(x = seq(0, 1, length = 10^(5)),
       y = with(prior_params, beta_density(x,
                                     mean = s_untested_mean,
                                     sd = s_untested_sd,
                                     bounds = s_untested_bounds))) %>%
   ggplot(aes(x=x, y = y)) +
  geom_line(alpha = .8) + 
  geom_ribbon(aes(x=x,ymin=0,ymax=y),
              fill="black",
              alpha=.6) +
  theme_c(legend.text=element_text(size = 10)) +
  viridis::scale_fill_viridis(discrete=TRUE,
                             option = "rocket", begin=.3,end=.8) +
  labs(x = "Value",
       y = "Probability Density", 
       title = latex2exp::TeX("Definition of Prior for $P(S_1|untested)$"),
       fill ='',
       subtitle = paste0("Mean: ", prior_params$s_untested_mean,
                         ", SD: ", prior_params$s_untested_sd))

```

# $P(S_0|  \text{test}_+, \text{untested})$

```{r}

tibble(x = seq(0, 1, length = 10^(5)),
       y = with(prior_params, beta_density(x,
                                     mean = p_s0_pos_mean,
                                     sd = p_s0_pos_sd,
                                     bounds = p_s0_pos_bounds))) %>%
   ggplot(aes(x=x, y = y)) +
  geom_line(alpha = .8) + 
  geom_ribbon(aes(x=x,ymin=0,ymax=y),
              fill="black",
              alpha=.6) +
  theme_c(legend.text=element_text(size = 10)) +
  viridis::scale_fill_viridis(discrete=TRUE,
                             option = "rocket", begin=.3,end=.8) +
  labs(x = "Value",
       y = "Probability Density", 
       title = latex2exp::TeX("Definition of Prior for $P(S_0|test_+,untested)$"),
       fill ='',
       subtitle = paste0("Mean: ", prior_params$p_s0_pos_mean,
                         ", SD: ", prior_params$p_s0_pos_sd))

```

# COVID-19 Trends and Impact Survey

```{r}

ctis_smoothed <- tar_read(ctis_smoothed,store = here("_targets"))

```

## $\beta$

```{r beta-smoothing, fig.cap="The ratio of the screening test positivity over the overall test positivity from the COVID-19 Trends and Impact Survey is taken to be the estimate of $\\beta$. We compare two approaches to smoothing: cubic spline smoothing with 2 knots (July 15th, 2021 and December 1st, 2021) to LOESS smoothing with a span of 0.33."}

ctis_smoothed  %>%
  filter(keep) %>%
  mutate(state=toupper(state)) %>%
  select(date,
         state, 
         imputed_beta,
         beta_estimate_smoothed, 
         beta_estimate_spline_smoothed) %>%
  pivot_longer(contains("beta")) %>%
  mutate(name = case_when(
    name == "beta_estimate_smoothed" ~ "LOESS smoothed",
    name == "beta_estimate_spline_smoothed" ~ "Spline smoothed",
    name == "imputed_beta" ~ "Survey Value"
  )) %>%
  ggplot(aes(x=date, y=value, color = name, 
             alpha = name, linewidth=name)) +
  geom_line() +
  facet_wrap(~state, ncol=4) +
  scale_alpha_manual(values=c("LOESS smoothed" = .9, 
                            "Spline smoothed" = .9, 
                              "Survey Value"=.3),
                     name='') +
  scale_linewidth_manual(values = c("LOESS smoothed" = 1.05, 
                            "Spline smoothed" = 1.05, 
                              "Survey Value"=.5)) +
  scale_color_manual(values=c("#3381FF", "#B58746", "#26900F"),
                     name ='') +
  guides(alpha="none",
         linewidth="none",
         color = guide_legend(override.aes = list(linewidth = 3,
                                                  alpha = c("LOESS smoothed" = .9, 
                            "Spline smoothed" = .9, 
                              "Survey Value"=.3)),
                            nrow=3)) +
  theme_c(legend.position="top") +
  ylim(0,1) +
  scale_x_date(date_breaks="3 months",
               date_labels = "%b %Y") + 
  labs(y = TeX("Survey Estimate of $\\beta$"),
       x= "",
       title = TeX("Comparing Approaches for Smoothing Survey Estimates of $\\beta$"))

```

## $\Pr(S_1 | \text{untested})$

```{r s-untested-smoothing, fig.cap="The percentage of the population experiencing COVID-19-like illness is taken to be the estimate of $\\Pr(S_1|\\text{untested}). The LOESS smoothed estimate with a span of 0.2 is shown in red."}

ctis_smoothed  %>%
  mutate(state=toupper(state)) %>%
  select(date,
         state, 
         contains("s_untested")) %>%
  pivot_longer(contains("s_untested")) %>%
  mutate(name = case_when(
    name == "s_untested_smoothed" ~ "LOESS smoothed",
    name == "imputed_s_untested" ~ "Survey Value"
  )) %>%
  ggplot(aes(x=date, y=value, color = name, 
             alpha = name, linewidth=name)) +
  geom_line() +
  facet_wrap(~state, ncol=5) +
  scale_alpha_manual(values=c("LOESS smoothed" = .9, 
                              "Survey Value"=.6),
                     name='') +
  scale_linewidth_manual(values = c("LOESS smoothed" = 1.02, 
                              "Survey Value"=.9)) +
  scale_color_manual(values=c("Survey Value" ="black", "LOESS smoothed"="darkred"),
                     name ='') +
  guides(alpha="none",
         linewidth="none",
         color = guide_legend(override.aes = list(
           linewidth = 3,
           alpha = c("LOESS smoothed" = .9, 
                     "Survey Value"=.3)),
                            nrow=3)) +
  theme_c(legend.position="top") +
  scale_x_date(date_breaks="3 months",
               date_labels = "%b %Y") + 
  labs(y = TeX("Survey Estimate of $Pr(S_1|untested)$"),
       x= "",
       title = TeX("Comparing Approaches for Smoothing Survey Estimates of $Pr(S_1|untested)$"))

```