### 2023-06-20 reports/state_results.Rmd
update state results

---
title: "State Results"
author: "Quinn White"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    df_print: paged
    code_folding: hide
    css: !expr here::here('css/template.css')
  html_document:
    df_print: paged
    toc-depth: 4
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = "../output_html")})
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE,warning=FALSE)

library(tidyverse)
library(targets)
library(here)
library(latex2exp)

```

```{r load-data}

# clear labels for versions
versions <- tibble(
  version = c("v1", "v2", "v3", "v4", "v5", "v6"),
  vlabel = factor(
    c( "$Priors\\,Do\\,Not\\,Vary\\,by\\,County\\,and\\,Date$", 
    "$\\beta$ Centered at Emp. Value",
    "$P(S_1|untested)$ Centered at Emp. Value",
    "$P(S_1|untested)$ and $\\beta$ Centered at Emp. Values",
        "$\\beta$ Centered at Emp. Value (Spline Smoothed)",
         "$P(S_1|untested)$ and $\\beta$ Centered at Emp. Values ($\\beta$ Spline Smoothed)"
    ),
    levels = c(
        "$Priors\\,Do\\,Not\\,Vary\\,by\\,County\\,and\\,Date$", 
        "$\\beta$ Centered at Emp. Value",
         "$P(S_1|untested)$ Centered at Emp. Value",
        "$P(S_1|untested)$ and $\\beta$ Centered at Emp. Values",
        "$\\beta$ Centered at Emp. Value (Spline Smoothed)",
         "$P(S_1|untested)$ and $\\beta$ Centered at Emp. Values ($\\beta$ Spline Smoothed)"

    ) )
)

# read state-level results
targets_dir <- here("_targets_6_19")

state_res <- tar_read(state_v1, 
                   store =targets_dir) %>% 
  mutate(version = "v1") %>%
  bind_rows(
    tar_read(state_v2,
             store =targets_dir)
  ) %>%
   bind_rows(
    tar_read(state_v3, 
             store =targets_dir)
  ) %>%
   bind_rows(
    tar_read(state_v4, 
             store =targets_dir)
   )
  # )%>%
  #  bind_rows(
  #   tar_read(state_v5, 
  #            store =targets_dir)
  # )%>%
  #  bind_rows(
  #   tar_read(state_v6, 
  #            store =targets_dir)
  # )

covidestim <- tar_read(covidestim_biweekly_state,
                        store =targets_dir)

dates <- readRDS(here("data/data_raw/date_to_biweek.RDS"))
codes <- read_csv(here('data/demographic/statecodes.csv'))

state_res <- state_res %>%
  rename(state=fips) %>%
  left_join(versions) %>%
  mutate(state=toupper(state)) %>%
  left_join(covidestim, relationship= "many-to-many") %>%
  left_join(dates, relationship = "many-to-many")

```

```{r theme_c}

theme_c <- function(...){ 
   # font <- "Helvetica"   #assign font family up front
    theme_bw() %+replace%    #replace elements we want to change

    theme(

      #text elements
      plot.title = element_text(             #title
                   size = 16,                #set font size
                   face = 'bold',            #bold typeface
                   hjust = .5,
                   vjust = 3),

      plot.subtitle = element_text(          #subtitle
                   size = 12,
                   hjust = .5,
                   face = 'italic',
                   vjust = 3),               #font size

      axis.title = element_text(             #axis titles
                   size = 12),               #font size

      axis.text = element_text(              #axis text
                   size = 9),
      legend.text = element_text(size = 12),
      # t, r, b, l
      plot.margin = unit(c(1,.5,.5,.5), "cm"),
      legend.position = "right",
      strip.text.x = element_text(size = 11, face = "bold", color="white"),
      strip.text.y = element_text(size = 11, face = "bold", color="white",angle=270),
      strip.background = element_rect(fill = "#3E3D3D")
      ) %+replace%
      theme(...)

}

```

# Summarizing Concordance with Covidestim

```{r,fig.width=11}

# corrected %>%
#   mutate(in_interval = case_when(
#     exp_cases_lb <= infections & infections <= exp_cases_ub ~ "In Interval",
#     exp_cases_lb  > infections  ~ "Below Interval",
#     exp_cases_ub < infections ~ "Above Interval"
#   )) %>%
#   filter(!is.na(in_interval)) %>%
#   group_by(in_interval, state, vlabel) %>%
#   summarize(n_per_county=n()) %>%
#   group_by(vlabel, in_interval) %>%
#   summarize(n_per_version = sum(n_per_county)) %>%
#   group_by(vlabel) %>%
#   mutate(total = sum(n_per_version)) %>%
#   ungroup() %>%
#   mutate(prop=n_per_version/total)

by_version <- state_res %>%
  select(-date) %>%
  distinct() %>% 
  mutate(in_interval = case_when(
    exp_cases_lb <= infections & infections <= exp_cases_ub ~ "In Interval",
    exp_cases_lb  > infections  ~ "Below Interval",
    exp_cases_ub < infections ~ "Above Interval"
  )) %>%
  filter(!is.na(in_interval)) %>%
  group_by(in_interval, vlabel) %>%
  summarize(n=n()) %>%
  group_by(vlabel) %>%
  mutate(total = sum(n)) %>%
  mutate(prop=n/total)

labels <- by_version %>%
  arrange(prop) %>%
  pull(vlabel) %>%
  as.character() %>%
  unique()

by_version %>%
  mutate(in_interval = factor(in_interval, 
                             levels = c("Above Interval",
                                        "In Interval",
                                        "Below Interval"
                                        ))) %>%
   ggplot(aes(x= fct_reorder(vlabel,prop,max), 
             fill = in_interval, y =prop)) + 
  geom_bar(stat="identity",position="stack") +
  theme_c() +
  coord_flip()+ scale_fill_manual(values=c("In Interval" = "#79D2AF",
                             "Above Interval" = "#D2AF79",
                      "Below Interval"="#7997D2"),
                       breaks = c("Below Interval",
                               "In Interval",
                               "Above Interval")) +
  labs(x="", 
       y= "Proportion",
       fill = "Covidestim Median:",
       title = "Proportion Where Covidestim Median\nWas Within, Above, or Below the Median",
       subtitle = "Including All States") +
  theme_c() +
#  viridis::scale_fill_viridis(option="mako", begin=.2, discrete=TRUE) +
  theme_c(legend.position="right",
          legend.title = element_text(face="bold", size = 15),
          legend.text= element_text( size = 15),
          legend.spacing.y = unit(3, 'pt'),
          axis.text.y = element_text(size = 17, hjust=1)) +
  theme(
          plot.subtitle = element_text(size=18, 
                                       face='italic', 
                                       margin=margin(4,0,4,0))) +
  scale_x_discrete(labels = (TeX(labels)))

```

# Ratio of Estimated to Observed During Omicron

```{r,results='hide', fig.height=11,fig.width=8}

subset <- state_res %>%
  filter(vlabel == "$Priors\\,Do\\,Not\\,Vary\\,by\\,County\\,and\\,Date$")

subset %>% 
  left_join(codes) %>%
  filter(biweek == 27) %>%
  pull(date) %>% 
  range()

subset %>% 
  left_join(codes) %>%
  filter(biweek == 27) %>%
  ggplot(aes(y =fct_reorder(state_name,exp_cases_median/positive),
             xmin = exp_cases_lb/positive,
             xmax = exp_cases_ub/positive)) +
  geom_errorbar() +
  # geom_point(aes(y=state_name,x=exp_cases_median/positive),
  #            color = "darkred", size=.8, alpha=.5) +
  theme_c() + 
  theme(axis.text.y=element_text(size=15),
        axis.text.x = element_text(size=13),
        axis.title = element_text(size = 18),
        plot.subtitle = element_text(size=14)) +
  labs(y="", x="Ratio of Estimated Infections to Observed",
       title ="Ratio of Estimated Infections to Observed Infections\nDuring Omicron",
       subtitle = "December 31, 2021 through January 14, 2022")

```