### 2023-06-24 writing/manuscript.Rmd
initialize writing rmds

---
title: "Approximating the Number of COVID-19 Infections in the United States Across 2021"
author: "Quinn White"
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    toc: false
    number_sections: false

---

\newpage

# Abstract

# Introduction

# Results

## County Level in Massachusetts

## State-level (1 version)

# Methods

## Data

### Massachusetts County Level
### State Level
### Survey Data

## Statistical Methods

### Probabilistic bias analysis

### Bayesian Melding

### Specification of Priors