### 2023-06-24 writing/supplement.Rmd

---
title: "Supplement"
author: "Quinn White"
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    toc: true
    number_sections: false

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
library(knitr)
library(here)

```

# Specification of the Priors

```{r ctis-by-state, fig.cap ="Since $\\beta$ represents the ratio of $\\Pr(\\text{test}_+|\\text{untested}, S_0)$ to $\\Pr(\\text{test}_+|\\text{tested})$, we estimate $\\beta$ from the COVID-19 Trends and Impact Survey as the ratio of the screening test positivity over the overall test positivity. Here, we consider the 2.5\\% and 97.5\\% percentiles and mean for each state. Most have a mean near $20%$."}

include_graphics(here("figures/emp-estimate-beta-dist-by-state.pdf"))

```

# Sensitivity Analyses