### 2023-06-26 reports/county_figures.Rmd
improve and add county-level figures

---
title: "County Figures"
author: "Quinn White"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    df_print: paged
    code_folding: hide
    css: !expr here::here('css/template.css')
  html_document:
    df_print: paged
    toc-depth: 4
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = "../output_html")})
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE,cache=TRUE, fig.width=10, fig.height=7)

library(tidyverse)
library(targets)
library(here)
library(latex2exp)
library(kableExtra)
library(scales)
library(cowplot)
```

```{r theme_c}

theme_c <- function(...){ 
   # font <- "Helvetica"   #assign font family up front
    theme_bw() %+replace%    #replace elements we want to change

    theme(

      #text elements
      plot.title = element_text(             #title
                   size = 16,                #set font size
                   face = 'bold',            #bold typeface
                   hjust = .5,
                   vjust = 3),

      plot.subtitle = element_text(          #subtitle
                   size = 12,
                   hjust = .5,
                   face = 'italic',
                   vjust = 3),               #font size

      axis.title = element_text(             #axis titles
                   size = 12),               #font size

      axis.text = element_text(              #axis text
                   size = 9),
      legend.text = element_text(size = 12),
      # t, r, b, l
      plot.margin = unit(c(1,.5,.5,.5), "cm"),
      legend.position = "right",
      strip.text.x = element_text(size = 11, face = "bold", color="white"),
      strip.text.y = element_text(size = 11, face = "bold", color="white",angle=270),
      strip.background = element_rect(fill = "#3E3D3D")
      ) %+replace%
      theme(...)

}

```

```{r load-data}

targets_dir <- here("_targets")

county <- tar_read(ma_v1, 
                   store =targets_dir) %>% 
  mutate(version = "v1") %>%
  bind_rows(
    tar_read(ma_v2,
             store =targets_dir)
  ) %>%
   bind_rows(
    tar_read(ma_v3, 
             store =targets_dir)
  ) %>%
   bind_rows(
    tar_read(ma_v4, 
             store =targets_dir)
  )%>%
   bind_rows(
    tar_read(ma_v5,
             store =targets_dir)
  )%>%
   bind_rows(
    tar_read(ma_v6,
             store =targets_dir)
   )%>%
   bind_rows(
    tar_read(ma_v7,
             store =targets_dir)
   )

waste <- tar_read(waste, store =targets_dir)
county_names <- read_tsv(here("data/demographic/county_fips.tsv")) %>%
  rename_with(.cols = everything(), .fn =tolower) %>%
  bind_rows()

covidestim <- tar_read(covidestim_biweekly_county,
                        store =targets_dir) %>%
  select(-c(date)) %>% 
  distinct()

dates <- readRDS(here("data/data_raw/date_to_biweek.RDS"))

```

```{r}

wjoined <- county %>%
  # set Nantucket, Duke fips to Nantucket since we have wastewater data
  # for Nantucket
  mutate(fips = ifelse(grepl("25019", fips), "25019", fips)) %>%
  inner_join(waste) %>%
  left_join(covidestim, relationship = "many-to-many") %>%
  left_join(dates,  relationship = "many-to-many") %>%
  group_by(biweek) %>%
  mutate(mindate = min(date),
         maxdate = max(date)) %>%
  ungroup()

counties_with_waste <- wjoined %>%
  group_by(fips) %>%
  mutate(obs_w_notna = sum(!is.na(mean_conc))) %>%
  filter(obs_w_notna != 0) %>%
  pull(fips) %>%
  unique()

versions <- tibble(
  version = c("v1", "v2", "v3", "v4", "v5", "v6", "v7"),
  vlabel = factor(
    c( "$Priors\\,Do\\,Not\\,Vary\\,by\\,State\\,and\\,Date$", 
    "$\\beta$ Centered at Emp. Value",
    "$P(S_1|untested)$ Centered at Emp. Value",
    "$P(S_1|untested)$ and $\\beta$ Centered at Emp. Values",
        "$\\beta$ Centered at Emp. Value (Spline Smoothed)",
         "$P(S_1|untested)$ and $\\beta$ Centered at Emp. Values ($\\beta$ Spline Smoothed)",
        "$\\,$CTIS Priors that Do Not Vary by Date$\\,$"
    ),
    levels = c(
        "$Priors\\,Do\\,Not\\,Vary\\,by\\,State\\,and\\,Date$", 
        "$\\beta$ Centered at Emp. Value",
         "$P(S_1|untested)$ Centered at Emp. Value",
        "$P(S_1|untested)$ and $\\beta$ Centered at Emp. Values",
        "$\\beta$ Centered at Emp. Value (Spline Smoothed)",
         "$P(S_1|untested)$ and $\\beta$ Centered at Emp. Values ($\\beta$ Spline Smoothed)",
        "$\\,$CTIS Priors that Do Not Vary by Date$\\,$"

    ) )
)

wjoined <- wjoined %>%
  left_join(versions)

w_data <- wjoined %>% 
  select(-date) %>%
  distinct() %>%
  filter(fips %in% counties_with_waste) %>%
  filter(!is.na(mean_conc)) %>%
  group_by(fips, version) %>%
  mutate(mean_conc_std = mean_conc/ max(mean_conc),
         exp_cases_median_std = exp_cases_median / max(exp_cases_median),
         exp_cases_lb_std = exp_cases_lb /max(exp_cases_median),
         exp_cases_ub_std = exp_cases_ub / max(exp_cases_median),
         infections_std  = infections/max(infections),
         positive_std = positive/max(positive)) %>%
  ungroup()

```

# Correlations

## Normalization Strategy 1: Standardization

```{r}

w_data <- wjoined %>% 
  select(-date) %>%
  distinct() %>%
  filter(fips %in% counties_with_waste) %>%
  filter(!is.na(mean_conc)) %>%
  group_by(fips, version) %>%
  mutate(
    mean_conc_std = (mean_conc - mean(mean_conc))/ 
           sd(mean_conc),
    exp_cases_median_std = (exp_cases_median -  mean(exp_cases_median))/
                                   sd(exp_cases_median),
         infections_std = (infections - mean(infections))/
      sd(infections),
         exp_cases_lb_std = (exp_cases_lb - mean(exp_cases_lb))/
      sd(exp_cases_lb),
        exp_cases_ub_std = (exp_cases_ub - mean(exp_cases_ub))/
      sd(exp_cases_ub),
         positive_std = (positive - mean(positive))
    /sd(positive))  %>%
  ungroup()

```

### PB Counts and Wastewater

```{r}

# correlations between PB counts and wastewater concentrations

correlations <- w_data %>%
  group_by(version) %>%
  summarize(Correlation = cor(exp_cases_median_std,mean_conc_std),
            vlabel =unique(vlabel)) %>%
  mutate(x=1.8,
         y=.1)

w_data %>%
  group_by(version) %>%
  ggplot(aes(x=exp_cases_median_std, y = mean_conc_std)) +
  geom_point(alpha=.5) +
  geom_linerange(aes(xmin = exp_cases_lb_std,
                     xmax=exp_cases_ub_std, 
                     y = mean_conc_std),
                 alpha = .3) +
  geom_text(
           aes(x=x,y=y,  
               label=paste0("Correlation: ", 
                                      round(Correlation,3))), 
                data= correlations ) +
  facet_wrap(~vlabel,
             ncol=2,
             labeller= as_labeller(TeX, 
                                   default = label_parsed)) +
  theme_c() +
  labs(x = "Probabilistic Bias Counts\n(Standardized by County)",
       y = "Mean Wastewater Concentration\n(Standardized by County)",
       title = "Correlations Between Probabilistic Bias Counts\nand Mean Wastewater Concentration")

```

```{r}
plt_wastewater_pb <- w_data %>%
  group_by(version) %>%
  ggplot(aes(x=exp_cases_median_std, y = mean_conc_std)) +
  geom_point(alpha=.5) +
  geom_linerange(aes(xmin = exp_cases_lb_std,
                     xmax=exp_cases_ub_std, 
                     y = mean_conc_std),
                 alpha = .3) +
  geom_text(
           aes(x=x,y=y,  label=paste0("Correlation: ", 
                                      round(Correlation,3))), 
                data= correlations ) +
  facet_wrap(~vlabel,
             ncol=1,
             labeller= as_labeller(TeX, 
                                   default = label_parsed)) +
  theme_c() +
  labs(x = "Probabilistic Bias Counts\n(Standardized by County)",
       y = "Mean Wastewater Concentration\n(Standardized by County)",
       title = "")

```

### PB Counts and Covidestim

```{r}

# correlations between PB counts and Covidestim estimates

covidestim_correlations <- w_data %>%
  filter(!is.na(infections_std)) %>%
  group_by(vlabel) %>%
  # missing values are handled by casewise deletion
  summarize(Correlation = cor(exp_cases_median_std,infections_std,
                              use = "complete.obs")) %>%
  mutate(x=2.8,
         y=.1)

w_data %>%
  ggplot(aes(x=exp_cases_median_std, y = infections_std)) +
  geom_point(alpha=.5) +
  geom_linerange(aes(xmin = exp_cases_lb_std,
                     xmax=exp_cases_ub_std, 
                     y = infections_std),
                 alpha = .3) +
  geom_text(
           aes(x=x,y=y,  label=paste0("Correlation: ", 
                                      round(Correlation,3))), 
                data= covidestim_correlations ) +
  facet_wrap(~vlabel,
              ncol=2,
             labeller= as_labeller(TeX, 
                                   default = label_parsed)) +
  theme_c() +
  labs(x = "Probabilistic Bias Counts\n(Standardized by County)",
       y = "Covidestim Estimates\n(Standardized by County)",
       title = "Correlations Between Probabilistic Bias Counts\nand Covidestim Estimates")

plt_covidestim_pb <- w_data %>%
  ggplot(aes(x=exp_cases_median_std, y = infections_std)) +
  geom_point(alpha=.5) +
  geom_linerange(aes(xmin = exp_cases_lb_std,
                     xmax=exp_cases_ub_std, 
                     y = infections_std),
                 alpha = .3) +
  geom_text(
           aes(x=x,y=y,  label=paste0("Correlation: ", 
                                      round(Correlation,3))), 
                data= covidestim_correlations ) +
  facet_wrap(~vlabel,
              ncol=1,
             labeller= as_labeller(TeX, 
                                   default = label_parsed)) +
  theme_c() +
  labs(x = "Probabilistic Bias Counts\n(Standardized by County)",
       y = "Covidestim Estimates\n(Standardized by County)",
       title = "")

```

### Combined Figure

```{r, fig.height=10, fig.width=12}

cowplot::plot_grid(plt_wastewater_pb, plt_covidestim_pb, ncol=2)

cor_covid <- covidestim_correlations %>% 
  select(vlabel, 
         `Correlation with Covidestim Estimates`=Correlation)

cor_waste <- correlations %>% 
  select(vlabel, 
         `Correlation with Wastewater Concentrations`=Correlation)

cor_waste %>%
  left_join(cor_covid) %>%
  mutate(vlabel = gsub(
    "$Priors\\,Do\\,Not\\,Vary\\,by\\,State\\,and\\,Date$", 
    "Priors Do Not Vary by State and Date",
    vlabel, fixed=TRUE),vlabel=gsub(" at","\nat", vlabel)) %>%
  arrange(desc( `Correlation with Wastewater Concentrations`)) %>%
  rename(Implementation = vlabel) %>%
  kbl() %>%
  kable_styling(full_width = TRUE)

```

### Covidestim and Wastewater

```{r}

covidestim_correlations <- w_data %>%
  filter(version=="v1") %>%
  # missing values are handled by casewise deletion
  summarize(Correlation = cor(mean_conc_std,infections_std,
                              use = "complete.obs")) %>%
  mutate(x=1,
         y=.1)

w_data %>%
  filter(version=="v1") %>%
  ggplot(aes(x=infections_std, y = mean_conc_std)) +
  geom_point(alpha=.5) +
  geom_text(
           aes(x=x,y=y,  label=paste0("Correlation: ", 
                                      round(Correlation,3))), 
                data= covidestim_correlations ) +
  theme_c() +
  labs(x = "Covidestim Estimates\n(Standardized by County)",
       y = "Wastewater Concentrations\n(Standardized by County)",
       title = "Correlations Between Wastewater Concentrations\nand Covidestim Estimates")

pb_correlations <- w_data %>%
  filter(version=="v1") %>%
  # missing values are handled by casewise deletion
  summarize(Correlation = cor(positive_std,mean_conc_std,
                              use = "complete.obs")) %>%
  mutate(x=1.8,
         y=.1)

```

### Compare to Wastewater Over Time

```{r,fig.width=12, fig.height=15}

library(scales)

pal <- c("#10BAC5", "#1B10C5", "#EFB719", "#900C3F")

names(pal) <- c(#"$observed$", 
                '$P(S_1|untested)$ Centered at Emp. Value',
                '$P(S_1|untested)$ and $\\beta$ Centered at Emp. Values',
               "$Priors\\,Do\\,Not\\,Vary\\,by\\,County\\,and\\,Date$",
               "$\\beta$ Centered at Emp. Value"
               )

w_data <- w_data %>%
  group_by(name) %>%
  mutate(adjust = 1/(max(mean_conc))*(exp_cases_median[which.max(mean_conc)])) %>%
  ungroup()

waste_df <- w_data %>%
  select(biweek, mean_conc, name, adjust, fips) %>%
  distinct() %>%
  left_join(dates,relationship = "many-to-many" ) %>%
  group_by(name, biweek) %>%
  # center date within 2-week interval
  summarize(date = min(date) + days(7),
            mean_conc =unique(mean_conc),
            adjust=unique(adjust),
            fips = unique(fips)) %>%
  mutate(name = gsub("County, MA", "", name))

pltlist <- w_data %>%
  mutate(name = gsub("County, MA", "", name)) %>%
  left_join(dates,relationship = "many-to-many" ) %>%
  # filter(version %in% c("v1","v3")) %>%
  filter(version=="v1") %>%
  filter(biweek>=6) %>%
  group_by(biweek) %>%
  mutate(mindate = min(date),
         maxdate=max(date)) %>%
  ungroup() %>% 
  group_by(name) %>%
  group_split() %>%
  map(~ {

    county_fips <- .x$fips %>% unique()
    adj <- .x$adjust %>% unique()

    .x %>% 
 #   filter(fips == county_fips  & date >= begin_date & date <= end_date) %>%
    ggplot() +
      geom_ribbon(aes(x = date, 
                 ymin = exp_cases_lb,
                 ymax = exp_cases_ub,
                 fill = vlabel),
                 alpha = .5) +
      geom_line(data = waste_df[which(waste_df$fips == county_fips),],
                aes(x = date, y =mean_conc*adj, 
                    color = 'Wastewater Effective Concentrations'),
               # color = "#DB4048",
                size = 1.1) +
      geom_linerange(aes(xmin = mindate, 
                         xmax = maxdate, 
                         y = infections,
                         color = 'Covidestim Infection Estimates')) + 
      geom_point(data =  waste_df[which(waste_df$fips == county_fips),],
                aes(x = date, y =mean_conc*adj ),
                color = "#DB4048",
                alpha = .5,
                size = 1.2) +
      scale_y_continuous(sec.axis=sec_axis(~.x/adj,
                                           name = 'Mean Wastewater Effective Concentration',
                                           labels=comma),
                         labels=comma
                    ) +
      # facet_grid(name~vlabel, 
      #            labeller=labeller(vlabel =as_labeller(
      #              TeX, default=label_parsed))) +
      facet_wrap(~name, ncol=3) + 
      scale_fill_manual(values = pal, labels = c('','','',''), 
                        name = "Probabilistic Bias Intervals") +
     theme_c(legend.position="none") +
      labs(y = "Value (Standardized by County)",
           title = "",
           x= "") +
      scale_x_date(date_labels = "%b %Y") +
    scale_color_manual(
      name = '', 
      values = c(
        'Covidestim Infection Estimates' = 'darkblue',
        'Wastewater Effective Concentrations' = '#DB4048')) +
    guides(color = guide_legend(override.aes = list(size = 12,
                                                    linewidth=3)),
           fill = guide_legend(override.aes =list(size = 6))) 
    })

legend <- cowplot::get_legend(
  pltlist[[1]] + theme(legend.position="top")
)

grid <- cowplot::plot_grid(plotlist=pltlist, ncol =3)

cowplot::plot_grid(legend, grid, nrow=2, rel_heights=c(.05,.95))

```

```{r, eval=FALSE,include=FALSE}

 w_data_std <- w_data %>%
  group_by(name, version) %>%
  mutate(exp_cases_lb = exp_cases_lb /max(exp_cases_median),
         exp_cases_ub = exp_cases_ub /max(exp_cases_median),
         exp_cases_median = exp_cases_median / max(exp_cases_median),
         mean_conc=mean_conc/max(mean_conc),
         infections=infections/max(infections)) %>%
  mutate(name = gsub("County, MA", "", name)) %>%
  left_join(dates,relationship = "many-to-many" ) %>%
  filter(version %in% c("v1","v3")) %>%
  filter(biweek>=6) %>%
  group_by(biweek) %>%
  mutate(mindate = min(date),
         maxdate=max(date)) %>%
  ungroup()

waste_df <- w_data_std %>%
  select(biweek, mean_conc, name, fips) %>%
  distinct() %>%
  left_join(dates,relationship = "many-to-many" ) %>%
  group_by(name, biweek) %>%
  # center date within 2-week interval
  summarize(date = min(date) + days(7),
            mean_conc =unique(mean_conc),
            fips = unique(fips)) %>%
  mutate(name = gsub("County, MA", "", name))

w_data_std %>% 
 #   filter(fips == county_fips  & date >= begin_date & date <= end_date) %>%
    ggplot() +
      geom_ribbon(aes(x = date, 
                 ymin = exp_cases_lb,
                 ymax = exp_cases_ub,
                 fill = vlabel),
                 alpha = .5) +
      geom_line(data = waste_df,
                aes(x = date, y =mean_conc, 
                    color = 'Wastewater Effective Concentrations'),
               # color = "#DB4048",
                size = 1.1) +
      geom_linerange(aes(xmin = mindate, 
                         xmax = maxdate, 
                         y = infections,
                         color = 'Covidestim Infection Estimates')) + 
      geom_point(data =  waste_df,
                aes(x = date, y =mean_conc),
                color = "#DB4048",
                alpha = .5,
                size = 1.2) +
   #   scale_y_continuous(sec.axis=sec_axis(~.x/adj)) +
      facet_grid(name~vlabel, 
                 labeller=labeller(vlabel =as_labeller(
                   TeX, default=label_parsed))) +
      scale_fill_manual(values = pal, labels = c('','','',''), 
                        name = "Probabilistic Bias Intervals") +
      theme_bw() +
     theme_c() +
      labs(y = "Value (Standardized by County)",
           title = "",
           x= "") +
      scale_x_date(date_labels = "%b %Y") +
    scale_color_manual(
      name = '', 
      values = c(
        'Covidestim Infection Estimates' = 'darkblue',
        'Wastewater Effective Concentrations' = '#DB4048')) +
    guides(color = guide_legend(override.aes = list(size = 12,
                                                    linewidth=3)),
           fill = guide_legend(override.aes =list(size = 6)))

```

## Normalization Strategy 2: By Values at Biweek Where Wastewater Concentration is Maximized

```{r}

w_data <- wjoined %>% 
  select(-date) %>%
  distinct() %>%
  filter(fips %in% counties_with_waste) %>%
  filter(!is.na(mean_conc)) %>%
  group_by(fips, version) %>%
  mutate(mean_conc_std = mean_conc/ max(mean_conc),
         max_conc = max(mean_conc),
         exp_cases_median_std = exp_cases_median / exp_cases_median[which.max(mean_conc)],
         exp_cases_lb_std = exp_cases_lb /exp_cases_median[which.max(mean_conc)],
         exp_cases_ub_std = exp_cases_ub / exp_cases_median[which.max(mean_conc)],
         infections_std  = infections/ infections[which.max(mean_conc)],
         positive_std = positive/max(positive)) %>%
  ungroup()

```

### PB Counts and Wastewater

```{r}

# correlations between PB counts and wastewater concentrations

correlations <- w_data %>%
  group_by(version) %>%
  summarize(Correlation = cor(exp_cases_median_std,mean_conc_std),
            vlabel =unique(vlabel)) %>%
  mutate(x=1.8,
         y=.1)

w_data %>%
  group_by(version) %>%
  ggplot(aes(x=exp_cases_median_std, y = mean_conc_std)) +
  geom_point(alpha=.5) +
  geom_linerange(aes(xmin = exp_cases_lb_std,
                     xmax=exp_cases_ub_std, 
                     y = mean_conc_std),
                 alpha = .3) +
  geom_text(
           aes(x=x,y=y,  label=paste0("Correlation: ", 
                                      round(Correlation,3))), 
                data= correlations ) +
  facet_wrap(~vlabel,
             ncol=2,
             labeller= as_labeller(TeX, 
                                   default = label_parsed)) +
  theme_c() +
  labs(x = "Probabilistic Bias Counts\n(Standardized by County)",
       y = "Mean Wastewater Concentration\n(Standardized by County)",
       title = "Correlations Between Probabilistic Bias Counts\nand Mean Wastewater Concentration")

plt_wastewater_pb <- w_data %>%
  group_by(version) %>%
  ggplot(aes(x=exp_cases_median_std, y = mean_conc_std)) +
  geom_point(alpha=.5) +
  geom_linerange(aes(xmin = exp_cases_lb_std,
                     xmax=exp_cases_ub_std, 
                     y = mean_conc_std),
                 alpha = .3) +
  geom_text(
           aes(x=x,y=y,  label=paste0("Correlation: ", 
                                      round(Correlation,3))), 
                data= correlations ) +
  facet_wrap(~vlabel,
             ncol=1,
             labeller= as_labeller(TeX, 
                                   default = label_parsed)) +
  theme_c() +
  labs(x = "Probabilistic Bias Counts\n(Standardized by County by County)",
       y = "Mean Wastewater Concentration\n(Standardized by County by County)",
       title = "")

```

### PB Counts and Covidestim

```{r}

# correlations between PB counts and Covidestim estimates

covidestim_correlations <- w_data %>%
  filter(!is.na(infections_std)) %>%
  group_by(vlabel) %>%
  # missing values are handled by casewise deletion
  summarize(Correlation = cor(exp_cases_median_std,infections_std,
                              use = "complete.obs")) %>%
  mutate(x=1.8,
         y=.1)

w_data %>%
  ggplot(aes(x=exp_cases_median_std, y = infections_std)) +
  geom_point(alpha=.5) +
  geom_linerange(aes(xmin = exp_cases_lb_std,
                     xmax=exp_cases_ub_std, 
                     y = infections_std),
                 alpha = .3) +
  geom_text(
           aes(x=x,y=y,  label=paste0("Correlation: ", 
                                      round(Correlation,3))), 
                data= covidestim_correlations ) +
  facet_wrap(~vlabel,
              ncol=2,
             labeller= as_labeller(TeX, 
                                   default = label_parsed)) +
  theme_c() +
  labs(x = "Probabilistic Bias Counts\n(Standardized by County by County)",
       y = "Covidestim Estimates\n(Standardized by County by County)",
       title = "Correlations Between Probabilistic Bias Counts\nand Covidestim Estimates")

plt_covidestim_pb <- w_data %>%
  ggplot(aes(x=exp_cases_median_std, y = infections_std)) +
  geom_point(alpha=.5) +
  geom_linerange(aes(xmin = exp_cases_lb_std,
                     xmax=exp_cases_ub_std, 
                     y = infections_std),
                 alpha = .3) +
  geom_text(
           aes(x=x,y=y,  label=paste0("Correlation: ", 
                                      round(Correlation,3))), 
                data= covidestim_correlations ) +
  facet_wrap(~vlabel,
              ncol=1,
             labeller= as_labeller(TeX, 
                                   default = label_parsed)) +
  theme_c() +
  labs(x = "Probabilistic Bias Counts\n(Standardized by County by County)",
       y = "Covidestim Estimates\n(Standardized by County by County)",
       title = "")

```

### Combined Figure

```{r, fig.height=10, fig.width=12}

cowplot::plot_grid(plt_wastewater_pb, plt_covidestim_pb, ncol=2)

cor_covid <- covidestim_correlations %>% 
  select(vlabel, 
         `Correlation with Covidestim Estimates`=Correlation)

cor_waste <- correlations %>% 
  select(vlabel, 
         `Correlation with Wastewater Concentrations`=Correlation)

cor_waste %>%
  left_join(cor_covid) %>%
  mutate(vlabel = gsub(
    "$Priors\\,Do\\,Not\\,Vary\\,by\\,County\\,and\\,Date$", 
    "Priors Do Not Vary by County and Date",
    vlabel, fixed=TRUE),vlabel=gsub(" at","\nat", vlabel)) %>%
  arrange(desc( `Correlation with Wastewater Concentrations`)) %>%
  rename(Implementation = vlabel) %>%
  kbl() %>%
  kable_styling(full_width = TRUE)

```

### Covidestim and Wastewater

```{r}

covidestim_correlations <- w_data %>%
  filter(version=="v1") %>%
  # missing values are handled by casewise deletion
  summarize(Correlation = cor(mean_conc_std,infections_std,
                              use = "complete.obs")) %>%
  mutate(x=1,
         y=.1)

w_data %>%
  filter(version=="v1") %>%
  ggplot(aes(x=infections_std, y = mean_conc_std)) +
  geom_point(alpha=.5) +
  geom_text(
           aes(x=x,y=y,  label=paste0("Correlation: ", 
                                      round(Correlation,3))), 
                data= covidestim_correlations ) +
  theme_c() +
  labs(x = "Covidestim Estimates\n(Standardized by County by County)",
       y = "Wastewater Concentrations\n(Standardized by County by County)",
       title = "Correlations Between Probabilistic Bias Counts\nand Covidestim Estimates")

pb_correlations <- w_data %>%
  filter(version=="v1") %>%
  # missing values are handled by casewise deletion
  summarize(Correlation = cor(positive_std,mean_conc_std,
                              use = "complete.obs")) %>%
  mutate(x=1.8,
         y=.1)

```

### Compare to Wastewater Over Time

```{r,fig.width=12, fig.height=15}

library(scales)

pal <- c("#10BAC5", "#1B10C5", "#EFB719", "#900C3F")

names(pal) <- c(#"$observed$", 
                '$P(S_1|untested)$ Centered at Emp. Value',
                '$P(S_1|untested)$ and $\\beta$ Centered at Emp. Values',
               "$Priors\\,Do\\,Not\\,Vary\\,by\\,County\\,and\\,Date$",
               "$\\beta$ Centered at Emp. Value"
               )

waste_df <- w_data %>%
  select(biweek, mean_conc_std, name) %>%
  distinct() %>%
  left_join(dates,relationship = "many-to-many" ) %>%
  group_by(name, biweek) %>%
  # center date within 2-week interval
  summarize(date = min(date) + days(7),
            mean_conc_std=unique(mean_conc_std)) %>%
  mutate(name = gsub("County, MA", "", name))

w_data%>%
  mutate(name = gsub("County, MA", "", name)) %>%
  left_join(dates,relationship = "many-to-many" ) %>%
  filter(version %in% c("v1","v3")) %>%
  filter(biweek>=6) %>%
  group_by(biweek) %>%
  mutate(mindate = min(date),
         maxdate=max(date)) %>%
  ungroup() %>% 
 #   filter(fips == county_fips  & date >= begin_date & date <= end_date) %>%
    ggplot() +
      geom_ribbon(aes(x = date, 
                 ymin = exp_cases_lb_std,
                 ymax = exp_cases_ub_std,
                 fill = vlabel),
                 alpha = .5) +
      geom_line(data = waste_df,
                aes(x = date, y =mean_conc_std, 
                    color = 'Wastewater Effective Concentrations'),
               # color = "#DB4048",
                size = 1.1) +
      geom_linerange(aes(xmin = mindate, 
                         xmax = maxdate, 
                         y = infections_std,
                         color = 'Covidestim Infection Estimates')) + 
      geom_point(data = waste_df,
                aes(x = date, y =mean_conc_std ),
                color = "#DB4048",
                alpha = .5,
                size = 1.2) +
      facet_grid(name~vlabel, 
                 labeller=labeller(vlabel =as_labeller(
                   TeX, default=label_parsed))) +
      scale_fill_manual(values = pal, labels = c('','','',''), 
                        name = "Probabilistic Bias Intervals") +
      theme_bw() +
     theme_c() +
      labs(y = "Value (Standardized by County by County)",
           title = "",
           x= "") +
      scale_x_date(date_labels = "%b %Y") +
    scale_color_manual(
      name = '', 
      values = c(
        'Covidestim Infection Estimates' = 'darkblue',
        'Wastewater Effective Concentrations' = '#DB4048')) +
    guides(color = guide_legend(override.aes = list(size = 12,
                                                    linewidth=3)),
           fill = guide_legend(override.aes =list(size = 6)))

```

## Normalization Strategy 3: By Maximum (Separately)

For each county, normalize the wastewater concentration by the maximum concentration for that county and the probabilistic bias counts by the maximum *median* estimated counts for that county (2.5th percentile and 97.5th percentile also standardized by the maximum *median* estimated counts for that county).

### PB Counts and Wastewater

```{r}

# correlations between PB counts and wastewater concentrations

correlations <- w_data %>%
  group_by(version) %>%
  summarize(Correlation = cor(exp_cases_median_std,mean_conc_std),
            vlabel =unique(vlabel)) %>%
  mutate(x=1.8,
         y=.1)

w_data %>%
  group_by(version) %>%
  ggplot(aes(x=exp_cases_median_std, y = mean_conc_std)) +
  geom_point(alpha=.5) +
  geom_linerange(aes(xmin = exp_cases_lb_std,
                     xmax=exp_cases_ub_std, 
                     y = mean_conc_std),
                 alpha = .3) +
  geom_text(
           aes(x=x,y=y,  label=paste0("Correlation: ", 
                                      round(Correlation,3))), 
                data= correlations ) +
  facet_wrap(~vlabel,
             ncol=2,
             labeller= as_labeller(TeX, 
                                   default = label_parsed)) +
  theme_c() +
  labs(x = "Probabilistic Bias Counts\n(Standardized by County)",
       y = "Mean Wastewater Concentration\n(Standardized by County)",
       title = "Correlations Between Probabilistic Bias Counts\nand Mean Wastewater Concentration")

plt_wastewater_pb <- w_data %>%
  group_by(version) %>%
  ggplot(aes(x=exp_cases_median_std, y = mean_conc_std)) +
  geom_point(alpha=.5) +
  geom_linerange(aes(xmin = exp_cases_lb_std,
                     xmax=exp_cases_ub_std, 
                     y = mean_conc_std),
                 alpha = .3) +
  geom_text(
           aes(x=x,y=y,  label=paste0("Correlation: ", 
                                      round(Correlation,3))), 
                data= correlations ) +
  facet_wrap(~vlabel,
             ncol=1,
             labeller= as_labeller(TeX, 
                                   default = label_parsed)) +
  theme_c() +
  labs(x = "Probabilistic Bias Counts\n(Standardized by County)",
       y = "Mean Wastewater Concentration\n(Standardized by County)",
       title = "")

```

### PB Counts and Covidestim

```{r}

# correlations between PB counts and Covidestim estimates

covidestim_correlations <- w_data %>%
  filter(!is.na(infections_std)) %>%
  group_by(vlabel) %>%
  # missing values are handled by casewise deletion
  summarize(Correlation = cor(exp_cases_median_std,infections_std,
                              use = "complete.obs")) %>%
  mutate(x=1.8,
         y=.1)

w_data %>%
  ggplot(aes(x=exp_cases_median_std, y = infections_std)) +
  geom_point(alpha=.5) +
  geom_linerange(aes(xmin = exp_cases_lb_std,
                     xmax=exp_cases_ub_std, 
                     y = infections_std),
                 alpha = .3) +
  geom_text(
           aes(x=x,y=y,  label=paste0("Correlation: ", 
                                      round(Correlation,3))), 
                data= covidestim_correlations ) +
  facet_wrap(~vlabel,
             ncol=2,
             labeller= as_labeller(TeX, 
                                   default = label_parsed)) +
  theme_c() +
  labs(x = "Probabilistic Bias Counts\n(Standardized by County)",
       y = "Covidestim Estimates\n(Standardized by County)",
       title = "Correlations Between Probabilistic Bias Counts\nand Covidestim Estimates")

plt_covidestim_pb <- w_data %>%
  ggplot(aes(x=exp_cases_median_std, y = infections_std)) +
  geom_point(alpha=.5) +
  geom_linerange(aes(xmin = exp_cases_lb_std,
                     xmax=exp_cases_ub_std, 
                     y = infections_std),
                 alpha = .3) +
  geom_text(
           aes(x=x,y=y,  label=paste0("Correlation: ", 
                                      round(Correlation,3))), 
                data= covidestim_correlations ) +
  facet_wrap(~vlabel,
              ncol=1,
             labeller= as_labeller(TeX, 
                                   default = label_parsed)) +
  theme_c() +
  labs(x = "Probabilistic Bias Counts\n(Standardized by County)",
       y = "Covidestim Estimates\n(Standardized by County)",
       title = "")
```

### Combined Figure

```{r, fig.height=10, fig.width=10}

cowplot::plot_grid(plt_wastewater_pb, plt_covidestim_pb, ncol=2)

```

### Covidestim and Wastewater

```{r}

covidestim_correlations <- w_data %>%
  filter(version=="v1") %>%
  # missing values are handled by casewise deletion
  summarize(Correlation = cor(mean_conc_std,infections_std,
                              use = "complete.obs")) %>%
  mutate(x=1,
         y=.1)

w_data %>%
  filter(version=="v1") %>%
  ggplot(aes(x=infections_std, y = mean_conc_std)) +
  geom_point(alpha=.5) +
  geom_text(
           aes(x=x,y=y,  label=paste0("Correlation: ", 
                                      round(Correlation,3))), 
                data= covidestim_correlations ) +
  theme_c() +
  labs(x = "Covidestim Estimates\n(Standardized by County)",
       y = "Wastewater Concentrations\n(Standardized by County)",
       title = "Correlations Between Probabilistic Bias Counts\nand Covidestim Estimates")

pb_correlations <- w_data %>%
  filter(version=="v1") %>%
  # missing values are handled by casewise deletion
  summarize(Correlation = cor(positive_std,mean_conc_std,
                              use = "complete.obs")) %>%
  mutate(x=1,
         y=.1)

```

# Simulation Intervals Over Time

```{r, fig.width=11, fig.height=8}

county_dat <- county %>%
  # set Nantucket, Duke fips to Nantucket since we have wastewater data
  # for Nantucket
  left_join(dates,  relationship = "many-to-many") %>%
  group_by(biweek) %>%
  mutate(mindate = min(date),
         maxdate = max(date)) %>%
  ungroup() %>%
  left_join(county_names, by=c('fips'='fips')) %>%
  mutate(name = ifelse(fips =="25007,25019", "Nantucket and Dukes", name))

county_dat %>%
  filter(version=="v1")  %>%
   ggplot() +
  geom_ribbon(aes(x = date, 
             ymin = exp_cases_lb,
             ymax = exp_cases_ub),
             alpha = .5,
             show.legend=FALSE,
             fill="#596281") +
   geom_linerange(aes(xmin = mindate,
                      xmax=maxdate,
                      y = positive,
                      color = 'obs'),
                  size=.5) +
  facet_wrap(~name, scales="free_y", ncol=3) +
  scale_y_continuous(labels = comma) +
  scale_x_date(date_breaks = "3 months", 
               date_labels = "%b %Y") +
  theme_c(
          legend.position = "top",
          ) +
  scale_fill_manual(values = pal)  +
  labs(x="",
       y = "Estimated Infections",
       title = "Simulation Intervals Over Time",
       subtitle = "Counties in Massachusetts") +
  scale_color_manual(values = c('obs' = 'red'), labels = 'Positive Tests',
                     name = '') +
  guides(color = guide_legend(override.aes = list(linewidth =2)))

```

## Ratio of Estimated Infections to Observed

```{r, fig.width=12, fig.height=9}

county_dat %>%
  filter(version=="v1")  %>%
  #filter(grepl("Nan",name)) %>%
   ggplot() +
  geom_ribbon(aes(x = date, 
             ymin = exp_cases_lb/positive,
             ymax = exp_cases_ub/positive),
             alpha = .5,
             show.legend=FALSE,
             fill="#596281") +
   # geom_linerange(aes(xmin = mindate,
   #                    xmax=maxdate,
   #                    y = positive,
   #                    color = 'obs')) +
  facet_wrap(~name,ncol=3) +
 # facet_grid(name~vlabel, 
 #                 labeller=labeller(vlabel =as_labeller(
 #                   TeX, default=label_parsed))) +
  scale_y_continuous(labels = comma) +
  scale_x_date(date_breaks = "3 months", 
               date_labels = "%b %Y") +
  theme_c(
          legend.position = "top") +
  scale_fill_manual(values = pal)  +
  labs(x="",
       y = "Estimated Infections",
       title = "Ratio of Estimated to Observed Over Time",
       subtitle = "Counties in Massachusetts") +
  scale_color_manual(values = c('obs' = 'red'), labels = 'Positive Tests',
                     name = '') +
  guides(color = guide_legend(override.aes = list(linewidth =2)))

```

# Trends in Total Tests and Positive Tests

Informs our understanding of the trends in the ratio of estimated to observed infections.

## Positive Tests and Test Positivity

```{r, fig.height=12, fig.width=12}

############################################
# POSITIVE TESTS AND TEST POSITIVITY
############################################
pltlist <- county_dat %>%filter(version=="v1") %>%
    mutate(testpos = positive/total) %>%
  group_by(across(-date)) %>%
  summarize(date=min(date) + days(7)) %>%
  group_by(fips) %>%
  mutate(adj = (1/max(testpos)) * max(positive)) %>%
  group_split() %>%
  map(~ {
    adj <-unique(.x$adj)
    .x %>%
      ggplot(aes(x=date))+
    geom_line(aes(y=positive, color = 'Positive Tests')) +
      geom_point(aes(y=positive, color = 'Positive Tests'),
              alpha=.5,
              size=.8) +
    geom_line(aes(y=testpos*adj, color='Test Positivity'),
              linetype=2) +
      geom_point(aes(y=testpos*adj, color='Test Positivity'),
              alpha=.5,
              size=.8) +
    #scale_linetype_discrete(breaks = c(1,1,2)) +
    scale_color_manual(values=c(
                                "Positive Tests"= "#279143",
                                "Test Positivity"= '#E38E4F')) +
    guides(color = guide_legend(override.aes = list(linetype = c(1,2),
                                                    linewidth=c(2,2)))) +
    theme_c() +
    theme(axis.title = element_text(size=10),
          axis.text =element_text(size=9),
          legend.position="none") +
    scale_y_continuous(labels=comma, sec.axis = sec_axis(~./adj , name = 'Test Positivity')) +
    labs(color ='',
         y="Positive Tests",
         x= "Date") +
      facet_wrap(~name)

  })

legend <- get_legend(
  pltlist[[1]] +
    theme(legend.position = "top",
          legend.text=element_text(size=16))
)

title_gg <- ggplot() + 
  labs(title = "Comparing Trends in Positive Tests and Test Positivity") + 
  theme(plot.title=element_text(face="bold",
                                hjust = .5, 
                                size = 18,
                                margin =margin(5,0,2,0)))

plt <- cowplot::plot_grid(plotlist=pltlist, ncol=3)

plot_grid(title_gg, legend, plt, rel_heights=c(.03, .03,.94), ncol=1)

```

## Total Tests and Positive Tests

```{r, fig.height=12, fig.width=12}

#####################################
# TOTAL TESTS AND POSITIVE TESTS
#####################################
pltlist <- county_dat %>%filter(version=="v1") %>%
    mutate(testpos = positive/total) %>%
  group_by(across(-date)) %>%
  summarize(date=min(date) + days(7)) %>%
  group_by(fips) %>%
  mutate(adj = (1/max(total)) * max(positive)) %>%
  group_split() %>%
  map(~ {
    adj <-unique(.x$adj)
    .x %>%
      ggplot(aes(x=date))+
    geom_line(aes(y=positive, color = 'Positive Tests')) +
      geom_point(aes(y=positive, color = 'Positive Tests'),
              alpha=.5,
              size=.8) +
    geom_line(aes(y=total*adj, color='Total Tests'),
              linetype=2) +
      geom_point(aes(y=total*adj, color='Total Tests'),
              alpha=.5,
              size=.8) +
    #scale_linetype_discrete(breaks = c(1,1,2)) +
    scale_color_manual(values=c(
                                "Positive Tests"= "#279143",
                                "Total Tests"= '#E38E4F')) +
    guides(color = guide_legend(override.aes = list(linetype = c(1,2)))) +
    theme_c() +
    theme(axis.title = element_text(size=11),
          axis.text =element_text(size=8),
          legend.position="none") +
    scale_y_continuous(labels=label_scientific(), 
                       n.breaks=4,
                       sec.axis = sec_axis(~./adj ,
                                           name = 'Total Tests',
                                           labels=label_scientific())) +
    labs(color ='',
         y="Positive Tests",
         x= "Date") +
      facet_wrap(~name)
})

legend <- cowplot::get_legend(
  pltlist[[1]] +
  theme(legend.position="top") +
    guides(color = guide_legend(keywidth=2,
                                override.aes = list(linewidth=c(2,2),
                                                    linetype =c(1,2))))
)

title_gg <- ggplot() + 
  labs(title = "Comparing Trends in Total Tests and Positive Tests") + 
  theme(plot.title=element_text(face="bold",
                                hjust = .5, 
                                size = 18,
                                margin =margin(5,0,2,0)))

plt <- cowplot::plot_grid(plotlist=pltlist, ncol=3)

cowplot::plot_grid(title_gg, legend, plt, rel_heights =c(.03, .03,.94),ncol=1)

```

## Percent Change in Total Tests and Positive Tests

```{r, fig.height=10, fig.width=12}

###################################################
# PERCENT CHANGE: TOTAL TESTS AND POSITIVE TESTS
###################################################

county_dat %>%
  filter(version=="v1") %>%
  filter(!grepl("Nantucket",name)) %>%
    mutate(testpos = positive/total) %>%
  group_by(across(-date)) %>%
  summarize(date=min(date) + days(7)) %>%
  group_by(fips) %>%
  arrange(biweek) %>%
  mutate(pct_change_positive = (positive - lag(positive, 1))/lag(positive, 1),
         pct_change_total = (total - lag(total, 1))/lag(total, 1)) %>%

   ggplot(aes(x=date))+
    geom_line(aes(y=pct_change_positive, color = 'Positive Tests')) +
      geom_point(aes(y=pct_change_positive, color = 'Positive Tests'),
              alpha=.5,
              size=.8) +
    geom_line(aes(y=pct_change_total, color='Total Tests'),
              linetype=2) +
      geom_point(aes(y=pct_change_total, color='Total Tests'),
              alpha=.5,
              size=.8) +
    #scale_linetype_discrete(breaks = c(1,1,2)) +
    scale_color_manual(values=c(
                                "Positive Tests"= "#279143",
                                "Total Tests"= '#E38E4F')) +
    guides(color = guide_legend(override.aes = list(linetype = c(1,2)))) +
    theme_c() +
    theme(axis.title = element_text(size=15),
          axis.text.x=element_text(size=9),
          legend.position="top") +
   # scale_y_continuous(labels=comma, sec.axis = sec_axis(~./adj , name = 'Test Positivity')) +
    labs(color ='',
         y="Percent Change",
         x= "Date",
         title = "Percent Change in Positive Tests and Total Tests") +
      facet_wrap(~name, ncol=3, scales="free_x") +
  scale_y_continuous(limits = c(-1,5.5),
                     labels=percent) +
  scale_x_date(date_breaks="3 months", date_labels="%b %Y")

```

# Concordance with Covidestim

```{r}

wjoined <- wjoined %>%
  mutate(name = gsub("County, MA","", name))

ma_concordance <- wjoined %>%
  filter(biweek >=6) %>%
  # covidestim does not report estimates for Nantucket 
  filter(!grepl("Nantucket", name )) %>%
  select(-date) %>%
  distinct() %>%
  select(exp_cases_lb, exp_cases_ub,
         name, biweek, vlabel, infections) %>%
  mutate(in_interval = case_when(
    exp_cases_lb <= infections & 
      infections <= exp_cases_ub ~ "In Interval",
    exp_cases_lb  > infections  ~ "Below Interval",
    exp_cases_ub < infections ~ "Above Interval"
  ) ) %>%
  group_by(in_interval, name, vlabel) %>%
  summarize(n_per_county=n()) %>%
  group_by(vlabel, in_interval) %>%
  summarize(n_per_version = sum(n_per_county)) %>%
  group_by(vlabel) %>%
  mutate(total = sum(n_per_version)) %>%
  ungroup() %>%
  mutate(prop=n_per_version/total)

ma_concordance_by_county <- wjoined %>%
  filter(biweek >=6) %>%
  # covidestim does not report estimates for Nantucket 
  filter(!grepl("Nantucket", name )) %>%
  select(-date) %>%
  mutate(name = gsub("$", "", 
                            name, fixed=TRUE)) %>%
  select(exp_cases_lb, exp_cases_ub, 
         name, biweek, vlabel, infections) %>%
   distinct() %>%
  mutate(in_interval = case_when(
    exp_cases_lb <= infections & 
      infections <= exp_cases_ub ~ "In Interval",
    exp_cases_lb  > infections  ~ "Below Interval",
    exp_cases_ub < infections ~ "Above Interval"
  ) ) %>%
  group_by(in_interval, name, vlabel) %>%
  summarize(n_per_county=n()) %>%
  group_by(name,vlabel) %>%
  mutate(total = sum(n_per_county),
         prop = n_per_county/total) %>%
  mutate(state="MA") %>%
  ungroup()

```

## Concordance by County

```{r}

##############################################
# stacked bar chart of concordance by county
##############################################
ma_concordance_by_county %>%
  filter(vlabel=="$P(S_1|untested)$ Centered at Emp. Value") %>%
  group_by(name) %>%
  mutate(m = prop[which(in_interval=="In Interval")]) %>%
  ungroup() %>%
  mutate(in_interval = factor(in_interval, levels = c("Above Interval",
                            "In Interval", "Below Interval"))) %>%
  ggplot(aes(x= fct_reorder(name,m), 
             fill = (in_interval), y =prop)) + 
  geom_bar(stat="identity",position="stack") +
  theme_c() +
  coord_flip()+
  scale_fill_manual(values=c("In Interval" = "#79D2AF",
                             "Above Interval" = "#D2AF79",
                      "Below Interval"="#7997D2")) +
  labs(x="", 
       y= "Proportion",
       fill = "Covidestim Median:",
       title = "Massachusetts: Proportion Where Covidestim Median\nWas Within, Above, or Below the Median",
       subtitle = TeX("For Implementation with $P(S_1|untested)$ Centered at Survey Value")) +
  theme_c() +
  theme(legend.position="right",
          legend.title = element_text(face="bold", size = 15),
          legend.text= element_text( size = 15),
          legend.spacing.y = unit(3, 'pt'),
          axis.text.y = element_text(size = 17, hjust=1))

```

## Concordance by Version

```{r, fig.width=13}

# 
in_interval <- ma_concordance %>%
  filter(in_interval == "In Interval") %>%
  mutate(n_interval = n_per_version) %>%
  select(vlabel, n_interval)

labels <- ma_concordance %>%
  filter(in_interval=="In Interval") %>%
  arrange((prop)) %>%
  pull(vlabel) %>%
  unique() %>%
  as.character()

#########################################################
# CONCORDANCE BY VERSION 
#########################################################
ma_concordance %>%
  left_join(in_interval) %>%
  mutate(in_interval=factor(in_interval, 
                            levels=c("Above Interval",
                                      "In Interval",
                                     "Below Interval"))) %>%
  ggplot(aes(y =fct_reorder(vlabel,n_interval),
             x = prop, 
             fill=in_interval)) +
  #facet_wrap(~fct_reorder((vlabel),m), ncol=1) +
  geom_bar(stat="identity",position="stack", color="darkgray", linewidth=.2) +
   theme_c(axis.text.y = element_text(size = 14, hjust=1),
          axis.title.x = element_text(size=14),
          axis.text.x=element_text(size=10),
          legend.position="right",
          legend.title = element_text(face="bold", size = 15),
          legend.text= element_text( size = 15),
          legend.spacing.y = unit(3, 'pt')) +
  scale_y_discrete(labels=unname(TeX(labels))) +
  labs(y ="",
       x="Proportion Containing the Covidestim Median",
       fill = "Covidestim Median:",
       title = "Where Covidestim Medians Fall\nRelative to Probabilistic Bias Intervals",
       subtitle = "By Implementation, in Massachusetts") +
  scale_fill_manual(values=c("Below Interval"="#7997D2",
                             "In Interval" = "#79D2AF",
                             "Above Interval" = "#D2AF79"),
                     breaks = c("Below Interval",
                               "In Interval",
                               "Above Interval")) +
  guides(fill=guide_legend(byrow=TRUE)) +
  scale_x_continuous(n.breaks = 7,
                     expand=c(0,0),
                     limits=c(-.05,1))

```

```{r, fig.width=12.3}

ma_concordance %>%
  group_by(vlabel) %>%
  mutate(total=sum(n_per_version)) %>%
  filter(in_interval == "In Interval") %>%
  mutate(n_interval = n_per_version) %>%
  ggplot(aes(y=fct_reorder(vlabel,prop), x=prop)) +
  geom_text(aes(label=round(prop,3), x= prop+.03), 
             angle=0) +
  geom_bar(stat="identity", fill="#596281") +
  scale_y_discrete(breaks= unique(ma_concordance$vlabel),
                   labels=function(x)TeX(x)) +
  scale_x_continuous(expand=c(0,0), 
                     limits=c(0,.9), n.breaks=7)  +
  theme_c(axis.text.y = element_text(size = 13, hjust=1),
          axis.title.x = element_text(size=14),
          axis.text.x=element_text(size=10)) +
  labs(y="",
       x="Proportion Containing the Covidestim Median",
       title="Proportion Containing the Covidestim Median",
       subtitle = "By Implementation")

```

# Summary of Comparison Between Implementations

Although the spline-smoothing improved the correlations with wastewater concentrations and the correlations with Covidestim estimates, it did not improve the proportion containing the Covidestim median.

The implementation centering $\Pr(S_1|\text{untested})$ at the survey value along with the implementation that does not vary the priors by state or date are the most correlated with Covidestim estimates, wastewater concentrations, and  have the highest proportion of county-biweeks containing the Covidestim median.