### 2023-06-27 reports/state_figures.Rmd

---
title: "State Results"
author: "Quinn White"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    df_print: paged
    code_folding: hide
    css: !expr here::here('css/template.css')
  html_document:
    df_print: paged
    toc-depth: 4
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = "../output_html")})
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE,warning=FALSE)

library(tidyverse)
library(targets)
library(here)
library(scales)
library(latex2exp)
library(cowplot)

```

```{r load-data, fig.width=10}

# clear labels for versions
versions <- tibble(
  version = c("v1", "v2", "v3", "v4", "v5", "v6", "v7"),
  vlabel = factor(
    c( "$Priors\\,Do\\,Not\\,Vary\\,by\\,State\\,and\\,Date$", 
    "$\\beta$ Centered at Emp. Value",
    "$P(S_1|untested)$ Centered at Emp. Value",
    "$P(S_1|untested)$ and $\\beta$ Centered at Emp. Values",
        "$\\beta$ Centered at Emp. Value (Spline Smoothed)",
         "$P(S_1|untested)$ and $\\beta$ Centered at Emp. Values ($\\beta$ Spline Smoothed)",
        "$\\,$CTIS Priors that Do Not Vary by State or Date$\\,$"
    ),
    levels = c(
        "$Priors\\,Do\\,Not\\,Vary\\,by\\,State\\,and\\,Date$", 
        "$\\beta$ Centered at Emp. Value",
         "$P(S_1|untested)$ Centered at Emp. Value",
        "$P(S_1|untested)$ and $\\beta$ Centered at Emp. Values",
        "$\\beta$ Centered at Emp. Value (Spline Smoothed)",
         "$P(S_1|untested)$ and $\\beta$ Centered at Emp. Values ($\\beta$ Spline Smoothed)",
        "$\\,$CTIS Priors that Do Not Vary by State or Date$\\,$"
    ) )
)

# read state-level results
targets_dir <- here("_targets_6_26")

state_res <- tar_read(state_v1, 
                   store =targets_dir) %>% 
  bind_rows(
    tar_read(state_v2,
             store =targets_dir)
  ) %>%
   bind_rows(
    tar_read(state_v3, 
             store =targets_dir)
  ) %>%
   bind_rows(
    tar_read(state_v4, 
             store =targets_dir)
   )%>%
   bind_rows(
    tar_read(state_v5,
             store =targets_dir)
  )%>%
   bind_rows(
    tar_read(state_v6,
             store =targets_dir)
  ) %>%
   bind_rows(
    tar_read(state_v7,
             store =targets_dir)
  )

covidestim <- tar_read(covidestim_biweekly_state,
                        store =targets_dir)

dates <- readRDS(here("data/data_raw/date_to_biweek.RDS"))
codes <- read_csv(here('data/demographic/statecodes.csv'))

state_res <- state_res %>%
  rename(state=fips) %>%
  left_join(versions) %>%
  mutate(state=toupper(state)) %>%
  left_join(covidestim, relationship= "many-to-many") %>%
  left_join(dates, relationship = "many-to-many")  %>%
  rename(name=state_name) %>%
  group_by(biweek) %>%
  mutate(mindate=min(date),
         maxdate=max(date)) %>%
  ungroup()

state_res <- state_res %>%
  filter(version %in% c("v2", "v5", "v6", "v7"))

```

```{r theme_c}

theme_c <- function(...){ 
   # font <- "Helvetica"   #assign font family up front
    theme_bw() %+replace%    #replace elements we want to change

    theme(

      #text elements
      plot.title = element_text(             #title
                   size = 16,                #set font size
                   face = 'bold',            #bold typeface
                   hjust = .5,
                   vjust = 3),

      plot.subtitle = element_text(          #subtitle
                   size = 12,
                   hjust = .5,
                   face = 'italic',
                   vjust = 3),               #font size

      axis.title = element_text(             #axis titles
                   size = 12),               #font size

      axis.text = element_text(              #axis text
                   size = 9),
      legend.text = element_text(size = 12),
      # t, r, b, l
      plot.margin = unit(c(1,.5,.5,.5), "cm"),
      legend.position = "right",
      strip.text.x = element_text(size = 11, face = "bold", color="white"),
      strip.text.y = element_text(size = 11, face = "bold", color="white",angle=270),
      strip.background = element_rect(fill = "#3E3D3D")
      ) %+replace%
      theme(...)

}

```

# Summarizing Concordance with Covidestim

```{r,fig.width=14}

# corrected %>%
#   mutate(in_interval = case_when(
#     exp_cases_lb <= infections & infections <= exp_cases_ub ~ "In Interval",
#     exp_cases_lb  > infections  ~ "Below Interval",
#     exp_cases_ub < infections ~ "Above Interval"
#   )) %>%
#   filter(!is.na(in_interval)) %>%
#   group_by(in_interval, state, vlabel) %>%
#   summarize(n_per_county=n()) %>%
#   group_by(vlabel, in_interval) %>%
#   summarize(n_per_version = sum(n_per_county)) %>%
#   group_by(vlabel) %>%
#   mutate(total = sum(n_per_version)) %>%
#   ungroup() %>%
#   mutate(prop=n_per_version/total)

by_version <- state_res %>%
  select(-date) %>%
  distinct() %>% 
  mutate(in_interval = case_when(
    exp_cases_lb <= infections & infections <= exp_cases_ub ~ "In Interval",
    exp_cases_lb  > infections  ~ "Below Interval",
    exp_cases_ub < infections ~ "Above Interval"
  )) %>%
  filter(!is.na(in_interval)) %>%
  group_by(in_interval, vlabel) %>%
  summarize(n=n()) %>%
  group_by(vlabel) %>%
  mutate(total = sum(n)) %>%
  mutate(prop=n/total)

labels <- by_version %>%
  arrange(prop) %>%
  pull(vlabel) %>%
  as.character() %>%
  unique()

by_version %>%
  mutate(in_interval = factor(in_interval, 
                             levels = c("Above Interval",
                                        "In Interval",
                                        "Below Interval"
                                        ))) %>%
   ggplot(aes(x= fct_reorder(vlabel,prop,max), 
             fill = in_interval, y =prop)) + 
  geom_bar(stat="identity",position="stack") +
  theme_c() +
  coord_flip()+ scale_fill_manual(values=c("In Interval" = "#79D2AF",
                             "Above Interval" = "#D2AF79",
                      "Below Interval"="#7997D2"),
                       breaks = c("Below Interval",
                               "In Interval",
                               "Above Interval")) +
  labs(x="", 
       y= "Proportion",
       fill = "Covidestim Median:",
       title = "Proportion of Simulation Intervals Where Covidestim Median\nWas Within, Above, or Below the Median",
       subtitle = "Including All States") +
  theme_c() +
  theme_c(legend.position="right",
          legend.title = element_text(face="bold", size = 15),
          legend.text= element_text( size = 15),
          legend.spacing.y = unit(3, 'pt'),
          axis.text.y = element_text(size = 17, hjust=1)) +
  theme(plot.title = element_text(size = 20),
          plot.subtitle = element_text(size=18, 
                                       face='italic', 
                                       margin=margin(4,0,4,0)),
        axis.text.x=element_text(size=12),
        axis.title = element_text(size=14)) +
  scale_x_discrete(labels = (TeX(labels)))

```

# Simulation Intervals Over Time

```{r}

all_versions <- state_res %>% 
  group_by(state) %>%
  summarize(state_n=n_distinct(version)) %>%
  filter(state_n == max(state_n)) %>%
  pull(state)
```

```{r, eval=FALSE}

state_res %>%
   filter(state %in% sample(unique(all_versions),5))  %>% 
   ggplot() +
  geom_ribbon(aes(x = date, 
             ymin = exp_cases_lb,
             ymax = exp_cases_ub),
             alpha = .8,
             show.legend=FALSE,
             fill="#596281") +
   geom_linerange(aes(xmin = mindate,
                      xmax=maxdate,
                      y = positive,
                      color = 'obs'),
                  size=.5) +
  facet_grid(name~vlabel, 
             labeller=labeller(vlabel =as_labeller(
               TeX, default=label_parsed)),
             scales="free_y") +
  scale_y_continuous(labels = comma) +
  scale_x_date(date_breaks = "3 months", 
               date_labels = "%b %Y") +
  theme_c(
          legend.position = "top",
          ) +
  scale_fill_manual(values = pal)  +
  labs(x="",
       y = "Estimated Infections",
       title = "Simulation Intervals Over Time",
       subtitle = "By State") +
  scale_color_manual(values = c('obs' = 'red'), labels = 'Positive Tests',
                     name = '') +
  guides(color = guide_legend(override.aes = list(linewidth =2)))

```

## Compare Versions

```{r}

subset <- state_res %>%
  filter(vlabel ==  "$\\,$CTIS Priors that Do Not Vary by State or Date$\\,$")

pal <- c("#247C90", "red")

names(pal) <- c(unique(subset$vlabel), "Covidestim")

state_res %>%
   filter(state %in% sample(unique(all_versions),5))  %>% 
  filter(version %in% c("v2", "v5", "v6", "v7")) %>%
   ggplot() +
  geom_ribbon(aes(x = date, 
             ymin = exp_cases_lb,
             ymax = exp_cases_ub,
             fill=vlabel),
             alpha = .8,
             show.legend=FALSE #,
           #  fill="#596281"
             ) +
   geom_linerange(aes(xmin = mindate,
                      xmax=maxdate,
                      y = positive,
                      color = 'obs'),
                  size=.5) +
  facet_grid(name~vlabel, 
             labeller=labeller(vlabel =as_labeller(
               TeX, default=label_parsed)),
             scales="free_y") +
  scale_y_continuous(labels = comma) +
  scale_x_date(date_breaks = "3 months", 
               date_labels = "%b %Y") +
  theme_c(
          legend.position = "top",
          ) +
  scale_fill_manual(values = pal)  +
  labs(x="",
       y = "Estimated Infections",
       title = "Simulation Intervals Over Time",
       subtitle = "By State") +
  scale_color_manual(values = c('obs' = 'red'), labels = 'Positive Tests',
                     name = '') +
  guides(color = guide_legend(override.aes = list(linewidth =2)))

```

## Compare States

```{r, fig.height=11, fig.height=9}

subset %>%
   ggplot() +
  geom_ribbon(aes(x = date, 
             ymin = exp_cases_lb,
             ymax = exp_cases_ub,
             fill = vlabel),
             alpha = .8,
             show.legend=FALSE) +
  geom_ribbon(aes(x = date,
                  fill = 'Covidestim',
                  ymin = infections.lo,
                  ymax = infections.hi),
              alpha = .6) +
  facet_wrap(~state, ncol=5, scales = "free_y",
              labeller= labeller(vlabel =as_labeller(TeX, default=label_parsed))) +
  scale_y_continuous(labels = comma) +
  scale_x_date(date_breaks = "3 months", 
               date_labels = "%b %Y") +
  theme_c(axis.text.x=element_text(size=6),
          axis.text.y = element_text(size=8),
          legend.position = "top") +
  scale_fill_manual(values =pal,
                    labels = TeX(names(pal)),
                    name='',
                    breaks="Covidestim")  +
  labs(y = "Estimated Infections",
       title = paste0("Estimated Infections by State")) +
  guides(color = guide_legend(override.aes = list(linewidth =2),
                              nrow=2,
                              byrow=TRUE))

ggsave(here('figures/intervals-and-covidestim-all-states.pdf'))

```

## Ratio of Estimated to Observed

```{r, fig.height=10,fig.width=12}

subset %>%
   ggplot() +
  geom_ribbon(aes(x = date, 
             ymin = exp_cases_lb/positive,
             ymax = exp_cases_ub/positive,
             fill = vlabel),
             alpha = .9,
             show.legend=FALSE,
             fill= "#247C90") +
  facet_wrap(~state, ncol=5, 
              labeller= labeller(vlabel =as_labeller(TeX, default=label_parsed))) +
  scale_y_continuous(labels = comma) +
  scale_x_date(date_breaks = "3 months", 
               date_labels = "%b %Y") +
  theme_c(axis.text.x=element_text(size=6),
          axis.text.y = element_text(size=8),
          legend.position = "top") +
  scale_fill_manual(values =pal,
                    labels = TeX(names(pal)),
                    name='',
                    breaks="Covidestim")  +
  labs(y = "Ratio of Estimated Infections to Observed",
       x="",
       title = paste0("Ratio of Estimated Infections to Observed by State")) +
  guides(color = guide_legend(override.aes = list(linewidth =2),
                              nrow=2,
                              byrow=TRUE))

ggsave(here('figures/ratio-estimated-to-observed-all-states.pdf'))

```

# Comparing States at Specific Two-week Intervals

## During Alpha Wave

```{r, fig.height=9,fig.width=8}

plt1 <- subset %>% 
  # left_join(codes) %>%
  filter(biweek == 7) %>%
  ggplot(aes(y =fct_reorder(name,exp_cases_median/positive),
             xmin = exp_cases_lb/positive,
             xmax = exp_cases_ub/positive)) +
  geom_errorbar() +
  # geom_point(aes(y=state_name,x=exp_cases_median/positive),
  #            color = "darkred", size=.8, alpha=.5) +
  theme_c() + 
  theme(axis.text.y=element_text(size=10),
        axis.text.x = element_text(size=11),
        axis.title = element_text(size = 17),
        plot.subtitle = element_text(size=14),
        plot.title = element_text(size=17, face="plain")) +
  labs(y="", x="Ratio of Estimated Infections\nto Observed",
       title ="During Two-week Period\nin the Alpha Wave",
       subtitle = "March 26, 2021 through April 8, 2021")

plt1

```

## During Delta Wave

```{r, fig.height=9,fig.width=8, results='hide'}

subset %>%
  mutate(ratio = exp_cases_median/positive) %>%
  group_by(biweek) %>%
  summarize(mean = mean(ratio),
            mindate=min(date),
            maxdate=max(date)) %>%
  arrange(desc(mean))

```

```{r, fig.height=9,fig.width=8}

plt2 <- subset %>% 
  # left_join(codes) %>%
  filter(biweek == 13) %>%
  ggplot(aes(y =fct_reorder(name,exp_cases_median/positive),
             xmin = exp_cases_lb/positive,
             xmax = exp_cases_ub/positive)) +
  geom_errorbar() +
  # geom_point(aes(y=state_name,x=exp_cases_median/positive),
  #            color = "darkred", size=.8, alpha=.5) +
  theme_c() + 
 theme(axis.text.y=element_text(size=10),
        axis.text.x = element_text(size=11),
        axis.title = element_text(size = 17),
        plot.subtitle = element_text(size=14),
        plot.title = element_text(size=17, face="plain")) +
  labs(y="", x="Ratio of Estimated Infections\nto Observed",
       title ="During Two-week Period\nin the Delta Wave",
       subtitle = "June 18, 2021 through July 1, 2021")

plt2

```

## During Omicron Wave

```{r,results='hide', fig.height=9,fig.width=8}

subset %>% 
  left_join(codes) %>%
  filter(biweek == 27) %>%
  pull(date) %>% 
  range()

subset %>% 
  left_join(codes) %>%
  filter(biweek == 13) %>%
  pull(date) %>% 
  range()

plt3 <- subset %>% 
  # left_join(codes) %>%
  filter(biweek == 27) %>%
  ggplot(aes(y =fct_reorder(name,exp_cases_median/positive),
             xmin = exp_cases_lb/positive,
             xmax = exp_cases_ub/positive)) +
  geom_errorbar() +
  # geom_point(aes(y=state_name,x=exp_cases_median/positive),
  #            color = "darkred", size=.8, alpha=.5) +
  theme_c() + 
  theme(axis.text.y=element_text(size=10),
        axis.text.x = element_text(size=11),
        axis.title = element_text(size = 17),
        plot.subtitle = element_text(size=14),
        plot.title = element_text(size=17, face="plain")) +
  labs(y="", x="Ratio of Estimated Infections\nto Observed",
       title ="During Two-week Period\nin the Omicron Wave",
       subtitle = "December 31, 2021 through January 14, 2022")

plt3

```

## All Waves Together

```{r, fig.height=10, fig.width=17}

plt <- plot_grid(plt1,plt2, plt3, nrow=1, align="none")

title <- ggplot() +
  labs(title = "Ratio of Estimated Infections to Observed Infections") +
  theme_c(plot.title=element_text(size=25))

plot_grid(title, plt, nrow=2, rel_heights= c(.05,.95))

```