### 2023-06-30 writing/supplement.Rmd
update supplement

---
title: "Supplement"
author: "Quinn White"
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    toc: true
    number_sections: false

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
library(knitr)
library(here)

```

# Specification of the Priors

```{r ctis-by-state, fig.cap ="Since $\\beta$ represents the ratio of $\\Pr(\\text{test}_+|\\text{untested}, S_0)$ to $\\Pr(\\text{test}_+|\\text{tested})$, we estimate $\\beta$ from the COVID-19 Trends and Impact Survey as the ratio of the screening test positivity over the overall test positivity. Here, we consider the 2.5\\% and 97.5\\% percentiles and mean for each state. Most have a mean near $20%$."}

include_graphics(here("figures/emp-estimate-beta-dist-by-state.pdf"))

```

# Details on Each Implementation

```{r covidestim-concordance-state, fig.cap="For each implementation, the proportions of simulation intervals where the Covidestim median was below, within, or above the interval are shown, across all states used for the implementation."}

include_graphics(here('figures/concordance-covidestim-state.pdf'))

```

```{r, fig.cap="The proportions of simulation intervals where the Covidestim median was below, within, or above the intervals are shown, considering all counties except Dukes/Nantucket, where Covidestim estimates were not produced."}

include_graphics(here('figures/concordance-covidestim-county.pdf'))

```

\newpage

# Full Set of Simulation Intervals by State

```{r sim-intervals-state}

include_graphics(here('figures/intervals-and-covidestim-all-states.pdf'))

```

## COVID-19 Trends and Impact Survey

# County Results by Implementation