### 2023-07-10 reports/priors.Rmd
update figures

---
title: "Priors"
author: "Quinn White"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    df_print: paged
    code_folding: hide
    css: !expr here::here('css/template.css')
  html_document:
    df_print: paged
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = "../output_html")})
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)

library(tidyverse)
library(targets)
library(here)
library(latex2exp)
library(viridis)

# read priors from target
prior_params <- tar_read(prior_params,
                         store = here("_targets"))
# load beta_density, gamma_density functions
source(here("R/03-base-functions.R"))

theme_c <- function(...){ 
   # font <- "Helvetica"   #assign font family up front
    theme_bw() %+replace%    #replace elements we want to change

    theme(

      #text elements
      plot.title = element_text(             #title
                   size = 16,                #set font size
                   face = 'bold',            #bold typeface
                   hjust = .5,
                   vjust = 3),

      plot.subtitle = element_text(          #subtitle
                   size = 12,
                   hjust = .5,
                   face = 'italic',
                   vjust = 3),               #font size

      axis.title = element_text(             #axis titles
                   size = 12),               #font size

      axis.text.x = element_text(              #axis text
                   size = 12),
      legend.text = element_text(size = 14),
      # t, r, b, l
      plot.margin = unit(c(1,.5,.5,.5), "cm"),
      legend.position = "right",
      strip.text.x = element_text(size = 16, face = "bold", color="white"),
      strip.background = element_rect(fill = "#3E3D3D")
      ) %+replace%
      theme(...)

}

```

# $\alpha$

```{r}

tibble(x = seq(0, 1.3, length = 10^(5)),
       y = with(prior_params, gamma_density(x,
                                     mean = alpha_mean,
                                     sd = alpha_sd,
                                     bounds = alpha_bounds))) %>%
   ggplot(aes(x=x, y = y)) +
  geom_line(alpha = .8) + 
  geom_ribbon(aes(x=x,ymin=0,ymax=y),
              fill="black",
              alpha=.6) +
  theme_c(legend.text=element_text(size = 10)) +
  viridis::scale_fill_viridis(discrete=TRUE,
                             option = "rocket", begin=.3,end=.8) +
  labs(x = "Value",
       y = "Probability Density", 
       title = latex2exp::TeX("Definition of Prior for $\\alpha$"),
       fill ='',
       subtitle = paste0("Mean: ", prior_params$alpha_mean,
                         ", SD: ", prior_params$alpha_sd))

```

# $\beta$

```{r}

tibble(x = seq(0, 1, length = 10^(5)),
       y = with(prior_params, beta_density(x,
                                     mean = beta_mean,
                                     sd = beta_sd,
                                     bounds = beta_bounds))) %>%
   ggplot(aes(x=x, y = y)) +
  geom_line(alpha = .8) + 
  geom_ribbon(aes(x=x,ymin=0,ymax=y),
              fill="black",
              alpha=.6) +
  theme_c(legend.text=element_text(size = 10)) +
  viridis::scale_fill_viridis(discrete=TRUE,
                             option = "rocket", begin=.3,end=.8) +
  labs(x = "Value",
       y = "Probability Density", 
       title = latex2exp::TeX("Definition of Prior for $\\beta$"),
       fill ='',
       subtitle = paste0("Mean: ", prior_params$beta_mean,
                         ", SD: ", prior_params$beta_sd))

```

# $P(S_1|\text{untested})$

```{r}

tibble(x = seq(0, 1, length = 10^(5)),
       y = with(prior_params, beta_density(x,
                                     mean = s_untested_mean,
                                     sd = s_untested_sd,
                                     bounds = s_untested_bounds))) %>%
   ggplot(aes(x=x, y = y)) +
  geom_line(alpha = .8) + 
  geom_ribbon(aes(x=x,ymin=0,ymax=y),
              fill="black",
              alpha=.6) +
  theme_c(legend.text=element_text(size = 10)) +
  viridis::scale_fill_viridis(discrete=TRUE,
                             option = "rocket", begin=.3,end=.8) +
  labs(x = "Value",
       y = "Probability Density", 
       title = latex2exp::TeX("Definition of Prior for $P(S_1|untested)$"),
       fill ='',
       subtitle = paste0("Mean: ", prior_params$s_untested_mean,
                         ", SD: ", prior_params$s_untested_sd))

tibble(x = seq(0, 1, length = 10^(5)),
       y = with(prior_params, beta_density(x,
                                     mean = s_untested_mean,
                                     sd = s_untested_sd,
                                     bounds = s_untested_bounds))) %>%
   ggplot(aes(x=x, y = y)) +
  geom_line(alpha = .8) + 
  geom_ribbon(aes(x=x,ymin=0,ymax=y),
              fill="black",
              alpha=.6) +
  theme_c(legend.text=element_text(size = 10)) +
  viridis::scale_fill_viridis(discrete=TRUE,
                             option = "rocket", begin=.3,end=.8) +
  labs(x = "Value",
       y = "Probability Density", 
       title = latex2exp::TeX("Definition of Prior for $P(S_1|untested)$"),
       fill ='',
       subtitle = paste0("Mean: ", prior_params$s_untested_mean,
                         ", SD: ", prior_params$s_untested_sd))

```

# $P(S_0|  \text{test}_+, \text{untested})$

```{r}

tibble(x = seq(0, 1, length = 10^(5)),
       y = with(prior_params, beta_density(x,
                                     mean = p_s0_pos_mean,
                                     sd = p_s0_pos_sd,
                                     bounds = p_s0_pos_bounds))) %>%
   ggplot(aes(x=x, y = y)) +
  geom_line(alpha = .8) + 
  geom_ribbon(aes(x=x,ymin=0,ymax=y),
              fill="black",
              alpha=.6) +
  theme_c(legend.text=element_text(size = 10)) +
  viridis::scale_fill_viridis(discrete=TRUE,
                             option = "rocket", begin=.3,end=.8) +
  labs(x = "Value",
       y = "Probability Density", 
       title = latex2exp::TeX("Definition of Prior for $P(S_0|test_+,untested)$"),
       fill ='',
       subtitle = paste0("Mean: ", prior_params$p_s0_pos_mean,
                         ", SD: ", prior_params$p_s0_pos_sd))

tibble(x = seq(0, 1, length = 10^(5)),
       y = with(prior_params, beta_density(x,
                                     mean = .55,
                                     sd = .12,
                                     bounds = NA))) %>%
   ggplot(aes(x=x, y = y)) +
  geom_line(alpha = .8) + 
  geom_ribbon(aes(x=x,ymin=0,ymax=y),
              fill="black",
              alpha=.6) +
  theme_c(legend.text=element_text(size = 10)) +
  viridis::scale_fill_viridis(discrete=TRUE,
                             option = "rocket", begin=.3,end=.8) +
  labs(x = "Value",
       y = "Probability Density", 
       title = latex2exp::TeX("Definition of Prior for $P(S_0|test_+,untested)$"),
       fill ='',
       subtitle = paste0("Mean: ", .55,
                         ", SD: ", .12)) +
  scale_x_continuous(n.breaks=11)

tibble(samp =sample_beta_density(1e4, mean = .55,
                                     sd = .12,
                                     bounds = NA))

qbeta(.025, shape1= get_beta_params( mu = .55,
                                     sd = .12)$a,
      shape2= get_beta_params( mu = .55,
                                     sd = .12)$b
      )

qbeta(.975, shape1= get_beta_params( mu = .55,
                                     sd = .12)$a,
      shape2= get_beta_params( mu = .55,
                                     sd = .12)$b
      )

```

# COVID-19 Trends and Impact Survey

```{r}

ctis_smoothed <- tar_read(ctis_smoothed,store = here("_targets"))

```

## $\beta$

```{r beta-smoothing, fig.cap="The ratio of the screening test positivity over the overall test positivity from the COVID-19 Trends and Impact Survey is taken to be the estimate of $\\beta$. We compare two approaches to smoothing: cubic spline smoothing with 2 knots (July 15th, 2021 and December 1st, 2021) to LOESS smoothing with a span of 0.33.", fig.width=12,fig.height=12}

ctis_smoothed  %>%
  filter(keep) %>%
  mutate(state=toupper(state)) %>%
  select(date,
         state, 
         imputed_beta,
         beta_estimate_smoothed, 
         beta_estimate_spline_smoothed) %>%
  pivot_longer(contains("beta")) %>%
  mutate(name = case_when(
    name == "beta_estimate_smoothed" ~ "LOESS smoothed",
    name == "beta_estimate_spline_smoothed" ~ "Spline smoothed",
    name == "imputed_beta" ~ "Survey Value"
  )) %>%
  ggplot(aes(x=date, y=value, color = name, 
             alpha = name, linewidth=name)) +
  geom_line() +
  facet_wrap(~state, ncol=4) +
  scale_alpha_manual(values=c("LOESS smoothed" = .9, 
                            "Spline smoothed" = .9, 
                              "Survey Value"=.3),
                     name='') +
  scale_linewidth_manual(values = c("LOESS smoothed" = 1.05, 
                            "Spline smoothed" = 1.05, 
                              "Survey Value"=.5)) +
  scale_color_manual(values=c("#3381FF", "#B58746", "#26900F"),
                     name ='') +
  guides(alpha="none",
         linewidth="none",
         color = guide_legend(override.aes = list(linewidth = 3,
                                                  alpha = c("LOESS smoothed" = .9, 
                            "Spline smoothed" = .9, 
                              "Survey Value"=.3)),
                            nrow=3)) +
  theme_c(legend.position="top") +
  ylim(0,1) +
  scale_x_date(date_breaks="3 months",
               date_labels = "%b %Y") + 
  labs(y = TeX("Survey Estimate of $\\beta$"),
       x= "",
       title = TeX("Comparing Approaches for Smoothing Survey Estimates of $\\beta$"))

```

## $\Pr(S_1 | \text{untested})$

```{r s-untested-smoothing, fig.cap="The percentage of the population experiencing COVID-19-like illness is taken to be the estimate of $\\Pr(S_1|\\text{untested}). The LOESS smoothed estimate with a span of 0.2 is shown in red.",fig.width=12,fig.height=16}

ctis_smoothed  %>%
  mutate(state=toupper(state)) %>%
  select(date,
         state, 
         contains("s_untested")) %>%
  pivot_longer(contains("s_untested")) %>%
  mutate(name = case_when(
    name == "s_untested_smoothed" ~ "LOESS smoothed",
    name == "imputed_s_untested" ~ "Survey Value"
  )) %>%
  ggplot(aes(x=date, y=value, color = name, 
             alpha = name, linewidth=name)) +
  geom_line() +
  facet_wrap(~state, ncol=4) +
  scale_alpha_manual(values=c("LOESS smoothed" = .9, 
                              "Survey Value"=.6),
                     name='') +
  scale_linewidth_manual(values = c("LOESS smoothed" = 1.02, 
                              "Survey Value"=.9)) +
  scale_color_manual(values=c("Survey Value" ="black", "LOESS smoothed"="darkred"),
                     name ='') +
  guides(alpha="none",
         linewidth="none",
         color = guide_legend(override.aes = list(
           linewidth = 3,
           alpha = c("LOESS smoothed" = .9, 
                     "Survey Value"=.3)),
                            nrow=3)) +
  theme_c(legend.position="top") +
  scale_x_date(date_breaks="3 months",
               date_labels = "%b %Y") + 
  labs(y = TeX("Survey Estimate of $Pr(S_1|untested)$"),
       x= "",
       title = TeX("Comparing Approaches for Smoothing Survey Estimates of $Pr(S_1|untested)$"))

```

## Distributions of $\beta$ by State

```{r}

# raw ctis data

ctis_raw <- readRDS(here('data/data_raw/ctis_all_states.RDS'))

states_keep <- ctis_smoothed  %>%
  filter(keep) %>% pull(state) %>% unique()

ctis <- ctis_raw %>%
  select(signal, state=geo_value, date, value) %>%
  pivot_wider(names_from=signal, values_from=value) %>%
  filter(state %in% states_keep) %>%
  mutate(beta=smoothed_wscreening_tested_positive_14d/
           smoothed_wtested_positive_14d,
         state=toupper(state))

```

```{r ctis-by-state, fig.cap ="Since $\\beta$ represents the ratio of $\\Pr(\\text{test}_+|\\text{untested}, S_0)$ to $\\Pr(\\text{test}_+|\\text{tested})$, we estimate $\\beta$ from the COVID-19 Trends and Impact Survey as the ratio of the screening test positivity over the overall test positivity. Here, we consider the 2.5\\% and 97.5\\% percentiles and mean for each state. Most have a mean near 0.2.", fig.width=9}

ctis %>%
  group_by(state) %>%
  summarize(median = median(beta, na.rm=TRUE),
            mean = mean(beta, na.rm=TRUE),
            q1 = quantile(beta, .025,  na.rm=TRUE),
            q2 = quantile(beta, .975,  na.rm=TRUE)) %>%
  ggplot(aes(x= fct_reorder(state, mean, .desc=TRUE),
             y = mean)) +
  geom_errorbar(aes(ymin=q1,ymax=q2),width=.2) +
  geom_point(color="darkred", size=2) +
  scale_y_continuous(n.breaks=10) +
  labs(y= TeX("Empirical Estimates of $\\beta$"),
       x="State",
       title = TeX("Empirical Estimates of $\\beta$ from the COVID-19 Trends and Impact Survey"),
       subtitle = "2.7% Percentile, 97.5% Percentile, and Median By State" ) +
  theme_c(axis.title=element_text(size=12))

ggsave(here("figures/emp-estimate-beta-dist-by-state.pdf"))

```

```{r}
#
# Logistic transformation of the Beta CDF.
#
f.beta <- function(alpha, beta, x, lower=0, upper=1) {
  p <- pbeta(x, alpha, beta)
 log(p/(1-p))
}
#
# Sums of squares.
#
delta <- function(fit, actual) sum((fit-actual)^2)
#
# The objective function handles the transformed parameters `theta` and
# uses `f.beta` and `delta` to fit the values and measure their discrepancies.
#
objective <- function(theta, x, prob, ...) {
  ab <- exp(theta) # Parameters are the *logs* of alpha and beta
  fit <- f.beta(ab[1], ab[2], x, ...)
  return (delta(fit, prob))
}
#
# Solve two problems.
#
par(mfrow=c(1,2))
alpha <- 15; beta <- 22 # The true parameters
for (x in list(c(1e-3, 2e-3), c(1/3, 2/3))) {
  x.p <- f.beta(alpha, beta, x)        # The correct values of the p_i
  start <- log(c(1e1, 1e1))            # A good guess is useful here
  sol <- nlm(objective, start, x=x, prob=x.p, lower=0, upper=1,
             typsize=c(1,1), fscale=1e-12, gradtol=1e-12)
  parms <- exp(sol$estimate)           # Estimates of alpha and beta
  #
  # Display the actual and estimated values.
  #
  print(rbind(Actual=c(alpha=alpha, beta=beta), Fit=parms))
  #
  # Plot the true and estimated CDFs.
  #      
  curve(pbeta(x, alpha, beta), 0, 1, n=1001, lwd=2)
  curve(pbeta(x, parms[1], parms[2]), n=1001, add=TRUE, col="Red")
  points(x, pbeta(x, alpha, beta))
}
```

```{r, fig.height=7}

ctis %>%
  group_by(state) %>%
  summarize(median = median(beta, na.rm=TRUE),
            mean = mean(beta, na.rm=TRUE),
            q1 = quantile(beta, .025,  na.rm=TRUE),
            q2 = quantile(beta, .975,  na.rm=TRUE)) %>%
  ggplot(aes(y=fct_reorder(state, mean, .desc=TRUE))) +
  ggridges::geom_density_ridges(
    aes(x=beta,
        y=fct_reorder(state, beta, .fun =median)),
    data=ctis,
    fill="#596281",
    quantile_lines=TRUE,
    quantiles = c(.025,.975),
    quantile_fun=function(beta,...) c(
                                    #  quantile(beta,.025, na.rm=TRUE),
                                      median(beta,na.rm=TRUE)
                                    #  quantile(beta,.975, na.rm=TRUE)
                                      )) +
  # geom_errorbar(aes( xmin=q1,xmax=q2),width=.2) +
  scale_x_continuous(n.breaks=10, limits=c(0,1)) +
  labs(x= TeX("Empirical Estimates of $\\beta$"),
       y="State",
       title = TeX("Empirical Estimates of $\\beta$ from the COVID-19 Trends and Impact Survey"),
       subtitle = "Density Estimate of Distribution and Median by State" ) +
  theme_c(axis.title=element_text(size=12),
          axis.text.x=element_text(size=11))

```

```{r}

ctis %>%
  ggplot(aes(x = beta, fill=state)) +
  geom_density(alpha=.6) +
  scale_fill_viridis(option="mako", discrete=TRUE) +
  theme_c()

```

```{r,include=FALSE}

ctis %>%
  ggplot(aes(x = beta, fill=state)) +
  geom_density() +
  scale_fill_viridis(option="mako", discrete=TRUE) +
  theme_c() +
  facet_wrap(~state)

```

```{r}

ctis %>%
  group_by(state) %>%
  summarize(mean=mean(beta,na.rm=TRUE)) %>%
  pull(mean) %>%
  mean()

ctis %>%
  group_by(state) %>%
  summarize(mean=mean(smoothed_wcli,na.rm=TRUE)) %>%
  pull(mean) %>%
  mean()

```

## Compare Priors

```{r}

source(here("R/05-state-analysis.R"))

beta_shape = get_shape(ctis_smoothed,
                             option="beta",
                       quiet=FALSE)

tibble(x = seq(0, 1, length = 10^(4)),
       original = with(prior_params, beta_density(x,
                                     mean = beta_mean,
                                     sd = beta_sd,
                                     bounds = beta_bounds)),
       ctis= dbeta(x, shape1=beta_shape[1], 
                   shape2=beta_shape[2])) %>%
  pivot_longer(c(original,ctis))%>%
   ggplot(aes(x=x, y = value, fill=name)) +
  geom_line(alpha = .8) + 
  geom_ribbon(aes(x=x,ymin=0, ymax=value,fill=name),
              alpha=.6) +
  theme_c(legend.text=element_text(size = 10)) +
  viridis::scale_fill_viridis(discrete=TRUE,
                             option = "rocket", begin=.3,end=.8) +
  labs(x = "Value",
       y = "Probability Density", 
       title = latex2exp::TeX("Definition of Prior for $\\beta$"),
       fill ='')

```

```{r}

s_untested_shape = get_shape(ctis_smoothed,
                             option="s_untested",
                       quiet=FALSE)

tibble(x = seq(0, 1, length = 10^(4)),
       original = with(prior_params, beta_density(x,
                                     mean = s_untested_mean,
                                     sd = s_untested_sd,
                                     bounds = s_untested_bounds)),
       ctis= dbeta(x, shape1=s_untested_shape[1], 
                   shape2=s_untested_shape[2])) %>%
  pivot_longer(c(original,ctis))%>%
   ggplot(aes(x=x, y = value, fill=name)) +
  geom_line(alpha = .8) + 
  geom_ribbon(aes(x=x,ymin=0, ymax=value,fill=name),
              alpha=.6) +
  theme_c(legend.text=element_text(size = 10)) +
  viridis::scale_fill_viridis(discrete=TRUE,
                             option = "rocket", 
                             begin=.3,
                             end=.8) +
  labs(x = "Value",
       y = "Probability Density", 
       title = latex2exp::TeX("Definition of Prior for $Pr(S_1|untested)$"),
       fill ='')

```

```{r,eval=FALSE}

ctis_smoothed <- tar_read(ctis_smoothed,
                          store=here("_targets"))

ctis_smoothed %>% left_join(dates) %>%
  group_by(biweek,state) %>%
  summarize(beta_est=mean(beta_est,na.rm=TRUE)) %>%
  ungroup() %>%
  summarize(q1 = quantile(beta_est,.025, na.rm=TRUE),
                median = median(beta_est, na.rm=TRUE),
                q2 = quantile(beta_est,.975, na.rm=TRUE),
                sd=sd(beta_est,na.rm=TRUE))

ctis_smoothed %>%
      summarize(q1 = quantile(
        beta_estimate_spline_smoothed,.025, na.rm=TRUE),
        median = median(beta_estimate_spline_smoothed, na.rm=TRUE),
        q2 = quantile(beta_estimate_spline_smoothed,.975, na.rm=TRUE),
        sd=sd(beta_estimate_spline_smoothed,na.rm=TRUE))

```

# Bayesian Melding

```{r}

library(patchwork)
library(truncdist)

source(here('R/04-melding.R'))

params <- prior_params
params$beta_shape1 <- beta_shape[1]
params$beta_shape2 <- beta_shape[2]
params$s_untested_shape1 <- s_untested_shape[1]
params$s_untested_shape2 <- s_untested_shape[2]
params$direct_params <- TRUE
# 
# prior_params <- list(
#     alpha_mean = .95,
#     alpha_sd = 0.08,
#     alpha_bounds = NA,
#    # alpha_bounds = c(.8,1),
#     beta_mean = .15,
#     beta_sd =.09,
#     beta_bounds = NA,
#   #  beta_bounds = c(0.002, 0.4),
#     s_untested_mean = .03,
#     s_untested_sd = .0225,
#   #  s_untested_bounds = c(0.0018, Inf),
#     s_untested_bounds = NA,
#     p_s0_pos_mean = .4,
#     p_s0_pos_sd = .1225,
#  #  p_s0_pos_bounds = NA,
#     p_s0_pos_bounds = c(.25, .7),
#     pre_nsamp = 1e4,
#     post_nsamp = 1e4)

melded <- do.call(get_melded, prior_params)

theta_wide <- with(prior_params,  
 theta <- tibble(x= seq(0,1.4, length = 200),
                 alpha = gamma_density(x,
                                       mean = alpha_mean,
                                       sd = alpha_sd,
                                       bounds = alpha_bounds),
                    # beta= beta_density(x,
                    #                           mean = beta_mean,
                    #                           sd = beta_sd,
                    #                           bounds = beta_bounds),
                    # P_S_untested = beta_density(x,
                    #                             mean = s_untested_mean,
                    #                             sd = s_untested_sd,
                    #                             bounds = s_untested_bounds),
                 P_S0_pos = beta_density(x,
                                         mean = p_s0_pos_mean,
                                         sd = p_s0_pos_sd,
                                         bounds = p_s0_pos_bounds)
                 )) %>%
  mutate(beta = dbeta(x, shape1= beta_shape[1], shape2=beta_shape[2]),
         P_S_untested = dbeta(x, shape1=s_untested_shape[1],
                              shape2=s_untested_shape[2] ))

theta_samp<- with(prior_params,  
 theta <- tibble(alpha = sample_gamma_density(pre_nsamp,
                                                mean = alpha_mean,
                                                sd = alpha_sd,
                                                bounds = alpha_bounds),
                    beta= rbeta(pre_nsamp,
                                shape1= beta_shape[1], 
                                shape2=beta_shape[2]),
                    P_S_untested = rbeta(pre_nsamp,
                                         shape1=s_untested_shape[1],
                                         shape2=s_untested_shape[2])) %>%
        mutate(phi_induced = est_P_A_testpos(P_S_untested = P_S_untested,
                                             alpha = alpha,
                                             beta=beta))) %>%
  select(phi_induced)

theta <- theta_wide %>% 
  pivot_longer(-c(x)) %>%
    mutate(name = case_when(
      name == "alpha" ~"$\\alpha$",
      name == "beta" ~"$\\beta$",
      name == "phi_induced" ~ "$M(\\theta) = Pr(S_0|test+,untested)$",
      name == "P_S_untested" ~ "$Pr(S_1|untested)$",
      name == "P_S0_pos" ~ "$Pr(S_0|test+,untested)$")
    ) %>%
    mutate(name = factor(name,
                         levels = c(
                           "$\\alpha$",
                           "$\\beta$",
                           "$Pr(S_1|untested)$",
                           "$Pr(S_0|test+,untested)$",
                           "$M(\\theta) = Pr(S_0|test+,untested)$")))

```

```{r, fig.width=9,fig.height=9}

melded_long <- reformat_melded(melded$post_melding,
                               melded$pre_melding,
                               pre_nsamp = prior_params$pre_nsamp,
                              p_s0_pos_mean = prior_params$p_s0_pos_mean,
                              p_s0_pos_sd = prior_params$p_s0_pos_sd,
                              p_s0_pos_bounds = c(.3,.7)) %>%
  filter(type=="After Melding") %>%
  mutate(name = gsub("$P(S_1|untested)$",
                     "$Pr(S_1|untested)$", name,fixed=TRUE),
         name = gsub("$P(S_0|test+,untested)$",
                     "$Pr(S_0|test+,untested)$", name, fixed=TRUE)) %>%
    mutate(name = factor(name,
                         levels = c(
                           "$\\alpha$",
                           "$\\beta$",
                           "$Pr(S_1|untested)$",
                           "$Pr(S_0|test+,untested)$")))

plt1 <- theta %>% 
  filter(name != "$M(\\theta) = Pr(S_0|test+,untested)$" & name !=  "$Pr(S_0|test+,untested)$") %>%
  ggplot() +
  geom_ribbon(aes(x=x, ymin=0, ymax= value, fill='Before Melding'),alpha=.7, color=NA)+
  geom_density(aes(x=value, fill = 'After Melding'), 
               color=NA,
               data = melded_long[melded_long$name !=
                                    "$Pr(S_0|test+,untested)$",]) +
  facet_wrap(~name,
             labeller=  as_labeller(TeX, default = label_parsed),
             ncol = 3, scales="free_y") +
  theme_c()+
  theme(legend.position="right",
          axis.text.x=element_text(size=6),
          axis.title.x=element_text(size=12),
          axis.title.y=element_text(size=12)) +
  theme(legend.position="none") +
  scale_x_continuous(n.breaks=5) +
  labs(x="Value")+
scale_fill_manual(values = c("Before Melding"= "#5670BF", "Induced" = "#B28542",
                                 "After Melding" = "#418F6A"))

theta_samp <- theta_samp %>% mutate(name ="$Pr(S_0|test+,untested)$" )

plt2 <- theta %>% 
  filter(name==  "$Pr(S_0|test+,untested)$") %>%
  ggplot() +
  geom_ribbon(aes(x=x, ymin=0, ymax= value, fill='Before Melding'),alpha=.7)+
  geom_density(aes(x=value, fill = 'After Melding'),
               alpha=.7,color=NA,
               data = melded_long[melded_long$name ==
                                    "$Pr(S_0|test+,untested)$",]) +
    geom_density(aes(x= phi_induced, fill="Induced"), 
                 color = NA,
                 data = theta_samp,
                 alpha =.7) +
  facet_wrap(~name,
             labeller=  as_labeller(TeX, default = label_parsed),
             ncol = 3, scales="free_y") +
  theme_c() +
  theme(legend.position="right",
          axis.text.x=element_text(size=6),
          axis.title.x=element_text(size=12),
          axis.title.y=element_text(size=12))+
  scale_x_continuous(n.breaks=5, limits =c(0,1)) +
  labs(x="Value",
       fill = "")+
    scale_fill_manual(values = c("Before Melding"= "#5670BF", "Induced" = "#B28542",
                                 "After Melding" = "#418F6A"))

plt1 / plt2

ggsave(here('figures/melding-priors.pdf'))

```